\documentclass[../main.tex]{subfiles}

\begin{document}

%% ======================================================================
%%  Section 6 – Applications using HDF5
%% ======================================================================
\section{Applications using HDF5}
\label{sec:applications}

Applications treat HDF5 as inert data, but HDF5 files can carry
behavior-relevant metadata: paths, external references, huge
declared shapes, filter/plugin requirements, serialized
application objects, and model-layer definitions. That creates
integration bugs even when the HDF5 library is behaving as
designed.

The HDF Group's security policy~\cite{hdf5-sec-pol} already attempts the right scope
distinction: core libhdf5, official tools, Remote Code Execution (RCE)
via parsing/API abuse, and supply-chain compromise are in scope;
third-party plugins and application-level usage are usually
triaged or referred to their maintainers.

\begingroup
\scriptsize
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.15}
\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.18\widetablewidth}p{0.18\widetablewidth}p{0.32\widetablewidth}p{0.28\widetablewidth}@{}}
\caption{Common application-level HDF5 failure families}
\label{tab:application-hdf5-failure-families}\\
\hline
\textbf{Family} & \textbf{Typical failure} & \textbf{Examples} & \textbf{Why it is common} \\
\hline
\endfirsthead

\caption[]{Common application-level HDF5 failure families, continued}\\
\hline
\textbf{Family} & \textbf{Typical failure} & \textbf{Examples} & \textbf{Why it is common} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

malformed file metadata reaches C memory code &
buffer overflow, over-read, UAF, double-free, null dereference, divide-by-zero &
many HDFGroup records, including 2024 and 2025 issues in object headers, datatypes, dataspaces, filters, heaps, free-space metadata, and cache images &
HDF5 is a complex binary metadata graph, not a flat byte container\\
application deserialization layered on HDF5 &
arbitrary code execution &
Keras CVE-2025-9905: crafted \code{.h5/.hdf5} model archives can execute code
because legacy \code{.h5} loading did not honor \code{safe_mode=True} for
\code{Lambda}-layer pickled code~\cite{cve-2025-9905}. &
ML frameworks often use HDF5 as a model container, then deserialize
framework-level objects from it. \\
external references and file path indirection &
arbitrary local file read, path traversal, data exfiltration &
Keras CVE-2026-1669: crafted \code{.keras} model file using HDF5 external
dataset references could read local files~\cite{cve-2026-1669}. &
HDF5 supports links and external storage mechanisms. External links store both
target filename and object path; external storage can resolve raw-data files
through prefixes and current working directory rules~\cite{hdf5-data-model}. \\
unbounded resource allocation &
memory exhaustion / interpreter crash &
Keras CVE-2026-0897: crafted \code{.keras} archive with
\code{model.weights.h5} declaring an extremely large dataset shape could exhaust
memory and crash Python~\cite{cve-2026-0897}. &
HDF5 metadata can declare logical shapes much larger than the actual useful
payload; application code may allocate based on declared shape before policy
checks. \\
dynamic plugin loading and search paths &
local privilege escalation / arbitrary code loading &
TensorFlow CVE-2026-2492: HDF5 plugin handling loaded plugins from an unsecured
location, enabling local privilege escalation~\cite{cve-2026-2492}. &
HDF5 supports dynamic plugin loading, and plugin paths can be affected by
defaults or \code{HDF5_PLUGIN_PATH}; dynamic loading can also be disabled via
\code{H5PLset_loading_state()} or
\code{HDF5_PLUGIN_PRELOAD=::}~\cite{hdf5-dynamic-plugins}. \\
vulnerable or stale embedded HDF5 copies &
inherited CVEs, unpatched parser behavior &
matio CVE-2021-36977 involved a heap overflow related to use of HDF5
1.12.0; Apple CVE-2021-31009 says multiple issues were addressed by removing
HDF5~\cite{cve-2021-36977,cve-2021-31009}. &
Downstream products vendor, statically link, or lag HDF5 releases;
vulnerability scanners then attribute HDF5 defects to the application. \\
helper-tool exposure &
crashes or code execution through inspection/conversion utilities &
\code{h5dump} appears in older and newer issues; CVE-2026-34734 describes a
heap-use-after-free in the \code{h5dump} helper utility triggered by a malicious
\code{.h5} file~\cite{cve-2026-34734}. &
Applications often call \code{h5dump}, \code{h5repack}, or converters as
``safe inspection'' tools on untrusted files. \\
\end{longtable}
\endgroup

\paragraph{The unifying root cause:} The repeated mistake is \textit{trust-boundary confusion}.
HDF5 sits at the boundary between a file and the process.
Downstream frameworks often add another boundary: model loading,
dataset ingestion, scientific workflow execution, browser upload,
cloud service processing, or plugin resolution. Many CVEs happen
because those layers collapse into one assumption: “This is just an HDF5 file, so reading it is safe.”
That assumption is false. An HDF5 file can influence:

\begin{itemize}
    \item memory allocation size
    \item dataset rank and shape
    \item chunk count and cache pressure
    \item filter/plugin loading
    \item external file lookup
    \item link traversal
    \item datatype conversion
    \item application-level model reconstruction
    \item tool output rendering
\end{itemize}

The common issue is \textbf{unsafe interpretation of HDF5-controlled
metadata at the application boundary.}

\subsection{Audit Findings and Mitigation Priorities}

Mitigation has to be layered. Patching \texttt{libhdf5}
alone will not fix Keras model deserialization, TensorFlow plugin
path handling, or application-level external-reference policy.

\subsubsection*{Priority 1: Define an “untrusted HDF5” profile}

HDF5 should offer, and downstreams should use, a strict profile
for any file received from users, networks, archives, model hubs,
object stores, email, web uploads, or third-party datasets.

Default behavior for that profile:
\begin{verbatim}
dynamic plugins disabled
external links disabled or sandboxed
external raw dataset storage disabled or sandboxed
VDS source traversal restricted
maximum rank enforced
maximum element count enforced
maximum allocation size enforced
maximum metadata bytes enforced
maximum chunk count enforced
maximum filter expansion ratio enforced
cycles and deep graph traversal rejected
unknown or legacy-dangerous messages rejected
    unless explicitly allowed    
\end{verbatim}

Today, pieces exist: plugin loading can be controlled, and all dynamic plugins
can be disabled with \code{H5PLset_loading_state(0)} or \code{HDF5_PLUGIN_PRELOAD=::}.
But this needs to become a cohesive security profile, not a set of hidden knobs.

\subsubsection*{Priority 2: Ship a preflight validator}

Create a validator that reads metadata only and returns a policy decision before the application allocates large arrays or deserializes model objects.

Example output:
\begin{minted}{json}
{
  "safe_for_untrusted_load": false,
  "reasons": [
    "external dataset reference outside sandbox",
    "dataset declares 8.7 TB logical size",
    "requires dynamic filter plugin",
    "contains legacy Keras Lambda layer metadata"
  ]
}    
\end{minted}

For core HDF5, this validator should reason about file-format invariants.
For downstreams, it should expose application hooks:

\begin{verbatim}
on_external_reference(path)
on_dataset_shape(path, dtype, shape, logical_bytes)
on_filter_required(filter_id)
on_link_traversal(source, target)
on_attribute(name, size)
\end{verbatim}

Keras and TensorFlow could then enforce model-specific policy before creating Python objects.

\subsubsection*{Priority 3: Disable dangerous HDF5 features by default in ML model loading}

ML loaders should not behave like general HDF5 viewers.

For Keras/TensorFlow-style use:
\begin{verbatim}
do not allow Lambda or pickled code from legacy .h5
do not allow external dataset references
do not allow dynamic HDF5 plugins
do not eagerly allocate declared tensors before checking total
    byte budget
do not resolve paths relative to current working
    directory
do not honor HDF5_PLUGIN_PATH in model loading
    contexts
\end{verbatim}   
This directly addresses the Keras arbitrary-code issue, the Keras arbitrary-file-read issue, the Keras memory-exhaustion
issue, and the TensorFlow plugin-search-path issue.

\subsubsection*{Priority 4: Make external references opt-in}

External links and external raw dataset storage are legitimate HDF5 features,
but they are unsafe defaults for hostile content.
HDF5 documentation states that external links store a target
filename and object path, and external dataset storage search
behavior can involve the current working directory, absolute
paths, relative paths, \code{${ORIGIN}}, and \code{HDF5_EXTFILE_PREFIX}.

Mitigation:
\begin{verbatim}
reject absolute external paths
reject path traversal
reject current-working-directory lookup
require explicit sandbox root
require explicit allowlist by extension or path
log every external resolution
never follow external references during metadata
    preview
\end{verbatim}

For services, the safest default is: \code{external references = off}.

\subsubsection*{Priority 5: Treat plugins as executable code}

HDF5 filter plugins, VOL connectors, and VFDs are not data.
They are code. The HDF5 docs state that plugins are shared
libraries and that the default path can be overwritten by
\code{HDF5_PLUGIN_PATH}. That is enough to create an
application-level vulnerability if a privileged process uses an
attacker-controlled path.

Mitigation:
\begin{verbatim}
clear HDF5_PLUGIN_PATH, HDF5_PLUGIN_PRELOAD,
    HDF5_EXTFILE_PREFIX in services
set plugin loading state explicitly at process start
allow plugins only from root/admin-owned directories
reject writable plugin directories
require hashes or signatures for plugins
never load plugins in setuid, service, notebook-hub,
    web-worker, or privileged contexts unless explicitly
    allowlisted
\end{verbatim}

Sandboxing plugins\ldots

\subsubsection*{Priority 6: Require downstream SBOM and version reporting}

Downstream projects should report:
\begin{verbatim}
HDF5 version
linkage type: static / dynamic / vendored / system
enabled filters
enabled VOL/VFD connectors
plugin search path policy
external-reference policy
safe-mode policy
\end{verbatim}

This solves a practical problem: a user seeing “HDF5 CVE” in a
scanner often cannot tell whether they are affected through
\code{libhdf5}, an official tool, a vendored copy, a Python wheel,
a model loader, or a plugin.

\subsubsection*{Priority 7: Split CVE classification into scope tags}

For HDF5-related CVE triage, use explicit tags:
\begin{verbatim}
HDF5-core-parser
HDF5-official-tool
HDF5-legacy-tool
HDF5-plugin-loading
HDF5-external-reference
HDF5-downstream-deserialization
HDF5-downstream-resource-policy
HDF5-vendored-library
not-HDF5
duplicate/rejected
\end{verbatim}

This avoids bad prioritization. A heap overflow in
\code{H5T__conv_struct_opt} is not the same class as Keras
ignoring \code{safe_mode} for legacy \code{.h5} files, even though both
mention HDF5.

\subsubsection*{A practical mitigation matrix for downstream maintainers}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}
\begin{longtable}{@{}p{0.25\widetablewidth}p{0.71\widetablewidth}@{}}
\caption{Minimum safe policies for downstream HDF5 use cases}
\label{tab:downstream-hdf5-safe-policies}\\
\hline
\textbf{Downstream use case} & \textbf{Minimum safe policy} \\
\hline
\endfirsthead

\caption[]{Minimum safe policies for downstream HDF5 use cases, continued}\\
\hline
\textbf{Downstream use case} & \textbf{Minimum safe policy} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

viewing unknown \code{.h5} files &
run in a sandbox; disable plugins; disable external references; cap metadata
and output size \\
loading ML models &
never load untrusted legacy \code{.h5}; disable Lambda/pickled code; validate
tensor shapes before allocation \\
cloud upload service &
preflight only; no plugin loading; no external file access; resource limits;
process isolation \\
HPC workflow &
separate trusted internal files from imported files; scrub environment; pin
plugin directories; use module-level HDF5 version floors \\
desktop scientific app &
prompt before following external references; ship patched HDF5; avoid writable
global plugin dirs \\
package/distribution &
publish SBOM; expose HDF5 version; rebuild promptly after HDF5 security
releases \\
library binding &
expose security controls in the binding API, not just through environment
variables \\
\end{longtable}
\endgroup

\paragraph{What The HDF Group can do to reduce downstream CVEs:}
The most useful move is not just more parser hardening.
It is to give downstreams safe controls that are easy to use.

Recommended deliverables:
\begin{verbatim}
H5Pset_security_profile(fapl, H5_SECURITY_PROFILE_UNTRUSTED)
H5Pset_external_reference_policy(fapl, H5_EXTREF_DENY
    | H5_EXTREF_SANDBOX)
H5Pset_plugin_policy(fapl, H5_PLUGIN_DENY
    | H5_PLUGIN_ALLOWLIST)
H5Pset_max_logical_dataset_bytes(fapl, n)
H5Pset_max_metadata_bytes(fapl, n)
H5Pset_max_object_count(fapl, n)
H5Pset_max_link_traversal_depth(fapl, n)
h5validate --profile untrusted --json file.h5
\end{verbatim}

Then publish downstream recipes for Keras, TensorFlow, h5py,
netCDF, MATLAB, R, Java, and cloud services.

\end{document}
