\documentclass[../main.tex]{subfiles}

\begin{document}

%% ======================================================================
%%  Section 3 – HDF5 Core Library and File Format
%% ======================================================================
\section{HDF5 Core Library and File Format}
\label{sec:core-lib-format}

\subsection{Accidents Waiting to Happen -- HDF5 Safety}

\begin{itemize}
    \item Disk full~\cite{forum-11714}
    \item File corruption~\cite{forum-file-corruption}
    \item Long-term data accessibility and interpretability, see section~\ref{sec:long-haul}
\end{itemize}

\subsection{Verify and Verify -- HDF5 Security}

\begin{itemize}
    \item Integer arithmetic
    \item File format parsing issues
    \item Malicious plugins, see section~\ref{sec:extensions}
\end{itemize}

A CVE history summary is included in Appendix~\ref{app:cve-history-summary}.
The list mixes true HDF5-library issues, removed-tool issues,
duplicate/rejected/not-HDF5 entries, and entries with no commit URL.
The main pattern is not “many unrelated bugs”; it is repeated failure to validate
file-derived structure sizes, offsets, counts, message lengths, continuation chunks,
datatype metadata, or filter parameters before using them in low-level C memory
operations. The 2025 entries alone show repeated heap overflows, use-after-free,
null dereference, memory leak, and resource-consumption defects across object
headers, file-space metadata, type conversion, filters, and metadata cache code.

The fix commentary confirms that several CVEs were symptoms of corrupted metadata
not being rejected early. (Examples: the file-space page-size fix added a sanity
check immediately after reading \code{page\_size} from the file; the
object-header-continuation fix rejects zero continuation chunk sizes;
the self-referential continuation-message fix checks expected versus actual
deserialized object-header chunks; and the cache-image fix added buffer-size
arguments, overflow checks, input validation, and cleanup.)

The major root-cause categories are tighter than the CVE count
suggests, and are summarized in Table~\ref{tab:hdf5-cve-root-cause-categories}.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.28\widetablewidth}p{0.68\widetablewidth}@{}}
\caption{Major root-cause categories in the HDF5 CVE history}
\label{tab:hdf5-cve-root-cause-categories}\\
\hline
\textbf{Root-cause category} & \textbf{What it means for HDF5} \\
\hline
\endfirsthead

\caption[]{Major root-cause categories in the HDF5 CVE history, continued}\\
\hline
\textbf{Root-cause category} & \textbf{What it means for HDF5} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

unchecked file-derived sizes / counts / offsets &
The dominant class. HDF5 metadata controls object layout, heap blocks, chunks,
dataspaces, filters, links, and type descriptors. When those fields are malformed,
C code later trusts them. \\
inconsistent internal bookkeeping &
Attribute tables, cache images, continuation chunks, and free lists often needed
separate ``allocated capacity'' versus ``actual count'' tracking. \\
decode/encode asymmetry &
Several issues happen because decode did not reject bad messages, then encode/flush
later assumed valid native structures. \\
arithmetic precondition failures &
Divide-by-zero and multiplication-overflow checks were missing in chunking, layout,
b-tree, filter, and dataspace paths. \\
unsafe cleanup/lifetime paths &
Use-after-free, double-free, leaks, and null dereferences show that malformed input
often breaks unwinding logic as much as the primary decode logic. \\
unbounded traversal, recursion, loop-progress, and resource-limit failures & 
External-reference/path-opening behavior is a policy boundary, not necessarily a parser memory-safety defect.\\
legacy tool attack surface &
\code{gif2h5}, HDF to GIF conversion, \code{h5dump}, and \code{h5repack} appear
repeatedly. Removal of \code{gif2h5} was a structural mitigation, not just a patch. \\
triage quality, duplicate/variant clustering, reachability, and fix-completeness metadata &
Several entries have \code{commit_url: null}, ``not filed against HDF5,'' ``cannot
reproduce,'' ``not applicable,'' or tentative confidence. Those should not be counted
as reviewed HDF5 fixes without qualification. \\
\end{longtable}
\endgroup

\subsection{HDF5 Privacy: Assets and Exposure}

\subsubsection{Introduction} 

Privacy risks in the HDF5 core library and file format are not accidental; as discussed
in the Introduction, they are systematic and arise from multiple factors. The design and
implementation of the HDF5 file format and library present one important source of such risks.
Additional risks emerge from scientific workflows, software ecosystems, interoperability
mechanisms, and evolving data usage contexts.

Before describing the assets that require protection in HDF5 and the mechanisms through
which privacy exposure may occur, it is important to emphasize several fundamental
characteristics of HDF5 and the audit approach adopted in this document.

\begin{itemize}
    \item \emph{HDF5 privacy primarily concerns accidental or unintended exposure of sensitive information
          rather than malicious attacks}. Malicious attacks and unauthorized access are primarily
          addressed through cybersecurity mechanisms and are outside the scope of this section.
    \item \emph{HDF5, by design, does not provide privacy protections}. When privacy is a concern, users and organizations
           are responsible for protecting data throughout the entire data lifecycle. In this document,
           the term data refers not only to individual HDF5 objects within a single file, but also to
           collections of HDF5 files and external data accessible through the HDF5 library,
           connectors, and related software components.
    \item \emph{HDF5 neither implements nor enforces privacy standards}. Furthermore, even when privacy
          standards and technical safeguards are in place, sensitive information may still be exposed
          through legitimate data processing, contextual interpretation, pattern discovery, inference,
          aggregation, and other analytical activities.
    \item \emph{HDF5 does not manage authentication, authorization, or filesystem protection}. However, privacy
          risks extend beyond traditional security concerns. Sensitive information may be exposed during
          legitimate data use by authorized users, even within highly secure computing environments.
    \item \emph{The risk of sensitive information exposure generally increases over the data lifecycle}.
          Sensitive information stored in HDF5 may be exposed during any stage of the lifecycle,
          including data creation, processing, storage, sharing, transfer, reuse, and archival.
    \item \emph{HDF5 extensibility mechanisms and associated software ecosystems may alter the privacy characteristics of data}.
          Features such as filters, Virtual Object Layer (VOL) connectors, Virtual File Drivers (VFDs), and
          external applications, including conversion and analysis tools, may introduce new privacy
          exposure surfaces or modify existing ones.
\end{itemize}

The remainder of this section focuses on the assets that require protection in HDF5, the mechanisms
through which sensitive information may be exposed, and potential approaches for mitigating privacy risks.

\subsubsection{HDF5 Assets and their Exposure}

The primary assets requiring protection are the scientific data themselves.
In HDF5, these include data stored in HDF5 objects, such as files, groups, datasets, datatypes,
and dataspaces, which represent experimental observations, measurements, simulations, images,
sensor outputs, and computational results.

Scientific data may be exposed through direct access using HDF5-based software,
including command-line utilities, graphical viewers, and programming interfaces such as Python.
However, privacy-sensitive information may also be inferred through examination of the 
low-level storage organization and auxiliary metadata embedded within the file to support 
portability, interoperability, and efficient data access. For example, string attributes 
visible through standard operating system utilities (e.g., the Unix strings command) may 
disclose dataset names, experimental identifiers, or other descriptive information. 
Similarly, inspection of dataset headers may reveal storage layout, chunk organization, 
compression methods, or data acquisition patterns (e.g., frame-based acquisition stored 
as chunked datasets). Examination of embedded strings may also disclose the use of 
external storage or references to datasets located in other HDF5 files.

It is important to distinguish between protecting access to data values and protecting information 
about data representation. Even when raw data are protected through mechanisms such as 
compression or encryption, information describing their storage organization, layout, or 
access methods may remain accessible and reveal sensitive contextual information about the 
underlying data or the processes that generated them.

\paragraph{Data Assets}

The primary assets requiring protection are the scientific data themselves. In HDF5, these 
include data stored in HDF5 objects, such as files, groups, datasets, datatypes, and dataspaces, 
which represent experimental observations, measurements, simulations, images, 
sensor outputs, and computational results.

Scientific data may be exposed through direct access using HDF5-based software, including 
command-line utilities, graphical viewers, and programming interfaces such as Python. 
However, privacy-sensitive information may also be inferred through examination of the low-level 
storage organization and auxiliary metadata embedded within the file to support portability, 
interoperability, and efficient data access. For example, string attributes visible through 
standard operating system utilities (e.g., the Unix strings command) may disclose dataset 
names, experimental identifiers, or other descriptive information. Similarly, inspection 
of dataset headers may reveal storage layout, chunk organization, compression methods, 
or data acquisition patterns (e.g., frame-based acquisition stored as chunked datasets). 
Examination of embedded strings may also disclose the use of external storage or references 
to datasets located in other HDF5 files.

It is important to distinguish between protecting access to data values and protecting 
information about data representation. Even when raw data are protected through mechanisms 
such as compression or encryption, information describing their storage organization, layout, 
or access methods may remain accessible and reveal sensitive contextual information about 
the underlying data or the processes that generated them.


\paragraph{Metadata Assets}

User-defined metadata often constitute one of the most privacy-sensitive asset classes in HDF5. 
The format was intentionally designed as a self-describing data model that allows extensive 
contextual information to be stored alongside scientific data. While HDF5 attributes are the most 
recognizable form of user-defined metadata, they represent only one category of descriptive 
information maintained within an HDF5 file.

User-defined metadata include all user-created HDF5 objects and object properties that organize, 
identify, or describe stored data. These include groups, datasets, their associated datatypes 
and dataspaces, committed datatypes, and attributes attached to HDF5 objects. 
Each HDF5 attribute is itself a structured metadata object with a name, datatype, and dataspace. 
Additional metadata assets include file and object names (link names), committed datatype names, 
compound datatype member names, enumeration labels, object comments, and information 
stored in the user block. Collectively, these metadata describe the logical organization, 
semantics, and relationships of the stored data and frequently contain experiment-specific, 
operational, or organizational information.

Like scientific data, metadata assets may be exposed not only through HDF5 applications but 
also through direct examination of the binary file format. Inspection of file structures or 
extraction of embedded strings without using HDF5 software may reveal object names, 
descriptive attributes, storage relationships, user block contents, and other contextual 
information that could disclose sensitive information about the data, their origin, 
or the organization that produced them.

\paragraph{Structural Metadata Assets}

Privacy-sensitive information may also be inferred from the structural characteristics of HDF5 files. 
Structural metadata assets include user-defined organizational structures, such as 
the hierarchical arrangement of objects, dataset storage layouts (e.g., contiguous, chunked, 
compact, or external storage), filter pipelines, and relationships established through links, 
object references, region references, or virtual datasets (VDS). In addition, the HDF5 library 
maintains internal structural metadata that describe and manage these user-defined structures, 
including chunk indexing schemes, B-tree organizations, and other implementation-specific data structures.

Even when raw data are inaccessible, for example, because they are encrypted, compressed, 
stored externally, or protected by file-system access controls, structural metadata may 
disclose information about the purpose of the data, experimental design, workflow organization, 
relationships among datasets, or assumptions made by the generating application. 
Naming hierarchies, grouping strategies, storage layouts, and reference relationships may reveal 
semantic information about scientific activities or operational processes. Likewise, 
implementation-specific structures, such as chunk indexing formats or other internal metadata, 
may disclose characteristics of the HDF5 library version or software implementation 
used to create or process the file.

Like scientific data and user-defined metadata, structural metadata may be exposed through 
HDF5 applications or by direct examination of the binary file format without using HDF5 software.


\paragraph{Derived and Reconstructed Information}

Privacy-sensitive information in HDF5 may also arise through inference or reconstruction 
from combinations of available data and metadata. Derived assets include relationships 
inferred from multiple HDF5 datasets, external links, object references, 
region references, virtual dataset (VDS) mappings, or information obtained by 
combining HDF5 data with external sources. 
These relationships may reveal information that is not explicitly represented within any single dataset or file.

For example, correlating quality-control datasets with experimental observations, combining 
temporal and spatial measurements, or reconstructing relationships across linked HDF5 files 
may disclose sensitive characteristics that are not apparent when individual datasets 
are examined in isolation. Consequently, privacy-sensitive assets extend beyond explicitly 
stored data and metadata to include information that can be inferred through aggregation, 
correlation, reconstruction, or other forms of analysis.

\paragraph{Workflow and Applications Artifacts}

Privacy-sensitive assets in HDF5 extend beyond data and metadata stored within HDF5 files 
to include artifacts created by applications, processing workflows, and the surrounding 
software ecosystem. These artifacts include command outputs (e.g., h5dump output), 
diagnostic logs, temporary files, metadata snapshots, cached intermediate datasets, 
conversion artifacts, and outputs generated by plugins or connectors 
(e.g., binary files produced by an HDF5 VOL connector for another scientific data format).

Examples include VFD SWMR metadata snapshots, cached intermediate results, diagnostic logs, 
and metadata generated during translation between non-self-describing data formats and HDF5. 
Such artifacts may expose scientific data, metadata, storage organization, workflow decisions, 
or implementation details beyond what is contained in the original HDF5 file. 
Because they are often created automatically, these artifacts may persist beyond their 
intended lifetime and remain unnoticed by users focused primarily on the scientific datasets themselves.

Scientific data processing workflows routinely enrich, restructure, annotate, and transform 
datasets to improve interoperability, reproducibility, usability, or analytical value. 
However, these processes may also introduce or preserve contextual information that was not intended for disclosure. 
Automatically generated metadata, intermediate processing products, and derived datasets may 
reveal experimental assumptions, workflow decisions, data provenance, software 
implementation details, or scientific semantics that were absent from the original data. 
Consequently, privacy exposure may arise not from malicious activity or unauthorized access, 
but as an unintended consequence of routine scientific data processing.


\subsection{HDF5 Tools}

Not strictly part of the core library, but the tools are an important part of the ecosystem and have their own set of issues.

\subsection{Audit Findings and Mitigation Priorities}

\subsubsection{Security and Safety}

The blunt conclusion: HDF5’s CVE history is mostly malformed-file
parsing pressure against a complex binary metadata machine.
The fix direction should not be “patch each crash.” It should be
systematic metadata validation at decode boundaries, explicit
invariants for each on-disk structure, fuzz corpora tied to invariants,
and hard separation between parser acceptance and internal object
construction.

\paragraph{Goal:} HDF5 should be able to parse any byte stream as hostile input
and either reject it cleanly or produce a validated internal object graph.
Anything else is security debt.

\paragraph{Implementation program:} define parser invariants first, harden decode
boundaries, then make fuzzing, CI gates, and release governance
enforce them.

The next priority is parser normalization: decode file bytes into a
validated intermediate form before constructing internal objects,
so malformed metadata cannot reach trusted C paths.

Our goal should be: \textit{make malformed HDF5 files boring}.
Every corrupted file should terminate in a controlled
\texttt{H5E\_BADVALUE}, \texttt{H5E\_OVERFLOW}, \texttt{H5E\_CORRUPT},
or \texttt{H5E\_RESOURCE} path. No segfaults. No assertions. No leaks.
No undefined behavior. No late-stage encode/flush crash from a bad
decode.

The current CVE list includes many issues in \texttt{H5FS},
\texttt{H5C}, \texttt{H5O}, \texttt{H5T}, \texttt{H5Z}, and
vector-copy paths, with repeated heap overflows, use-after-free,
null dereference, and resource-consumption outcomes.
HDF5’s own format documentation explains why this happens: a file is a
graph of groups, datasets, and named datatypes; on disk it is
represented through lower-level structures such as the superblock,
B-tree nodes, object headers, global heaps, local heaps, and
free-space metadata. That means the library is parsing a dense graph
of mutually referential, \textit{attacker-controlled} metadata.

\subsubsection*{Priority 0: Stop closing CVEs as “fixed” unless the class is killed}

Make this rule immediate:


\textit{If the fix is not a structural change that eliminates the root cause, then it is not a fix.}

For every new CVE or crash:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.30\linewidth}p{0.66\linewidth}@{}}
\caption{Required artifacts for closing HDF5 CVEs or crash reports}
\label{tab:cve-closure-artifacts}\\
\hline
\textbf{Required artifact} & \textbf{Purpose} \\
\hline
\endfirsthead

\caption[]{Required artifacts for closing HDF5 CVEs or crash reports, continued}\\
\hline
\textbf{Required artifact} & \textbf{Purpose} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

crashing input & becomes a permanent corpus seed \\
minimized input & used in quick PR regression \\
root-cause category & maps to a known invariant class \\
invariant update & prevents sibling bugs \\
sanitizer regression & proves no crash, no leak, no UAF \\
negative test & checks controlled failure code \\
reviewer checklist update & prevents reintroduction \\
\end{longtable}
\endgroup

This is the first cultural shift. A patch that only adds a local \texttt{if}
is tolerated for embargoed fixes, but the issue remains open as a
\textbf{class bug} until the invariant framework catches sibling cases.

Recent commits show the right direction, but too locally. One fix
added a sanity check immediately after reading file-space page size
from the file and used the same fuzzer to verify that the previous
assertion became a corrupted-file error.~\cite{commit_804f3ba}
Another hardened \texttt{H5C\_\_reconstruct\_cache\_entry()} by adding
buffer-size tracking, overflow checks, input validation, and safe cleanup.~\cite{commit_3914bb7}
Those are exactly the moves to systematize.

\subsubsection*{Priority 1: Build an invariant registry for the file format}

Create a versioned document and machine-readable metadata file, for example:

\begin{verbatim}
security/invariants/
  superblock.yml
  object_header.yml
  dataspace.yml
  datatype.yml
  layout.yml
  filters.yml
  heaps.yml
  free_space.yml
  links.yml
  cache_image.yml
\end{verbatim}

Each entry should define:

\begin{minted}{yaml}
id: H5O-CONT-001
component: object_header
structure: continuation_message
condition: continuation_chunk_size > 0
failure: H5E_CORRUPT
checked_at:
  - H5O__cont_decode
  - H5O__chunk_deserialize
fuzz_targets:
  - fuzz_object_header
cves:
  - CVE-2025-6856
  - CVE-2025-6816
\end{minted}

Start with these invariant families.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.16\widetablewidth}p{0.80\widetablewidth}@{}}
\caption{Invariant families to enforce first}
\label{tab:invariant-families-first}\\
\hline
\textbf{Area} & \textbf{Invariants to enforce first} \\
\hline
\endfirsthead

\caption[]{Invariant families to enforce first, continued}\\
\hline
\textbf{Area} & \textbf{Invariants to enforce first} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

superblock &
offset size and length size are valid; base address, EOF address, driver info
address, and root object address are inside the file or explicitly undefined;
EOF does not wrap \\
object headers &
message length fits inside chunk; continuation chunks are nonzero; continuations
do not self-reference; continuations are acyclic; message flags match message
type; unknown messages cannot be cast into known message structures \\
free-space metadata &
page size is nonzero and below max; section size fits page/file bounds;
serialized section count matches decoded section count \\
heaps &
heap object offset and size are inside heap; free-list blocks do not overlap;
free-block list is acyclic; no allocation uses unchecked \code{count * size} \\
dataspaces &
rank is bounded; dimension product is checked; selected point count is checked;
hyperslab stride, count, block, and start do not overflow \\
datatypes &
storage size is nonzero; precision fits storage size; member offsets fit
compound size; string length is bounded; vlen/ref types cannot escape validation \\
layouts &
compact data size equals \code{dataspace elements * datatype size}; chunk
dimensions are nonzero; chunk product is checked; chunk index references are
inside file bounds \\
filters &
parameter count is bounded; decompressed size is bounded; scale-offset precision
is validated; filter output ratio cannot exceed configured resource limits \\
cache image &
cache-entry count, address, size, and alignment are bounded; decoding consumes
exactly the expected number of bytes; cleanup is safe after partial decode \\
tools &
string rendering is length-bounded; conversion tools never trust input
dimensions; tool-only parsers use the same checked arithmetic wrappers as the
library \\
\end{longtable}
\endgroup

This registry becomes the source of truth for code review, tests, fuzzing, and documentation.

\subsubsection*{Priority 2: Introduce a two-phase parser boundary}

Right now, too many paths appear to do this:
\begin{verbatim}
file bytes -> native internal object
           -> later code assumes validity
\end{verbatim}

Change the architecture to:
\begin{verbatim}
file bytes -> bounded raw decode
           -> semantic validation
           -> native object construction
\end{verbatim}

Concrete rule: \textit{No public or internal runtime object is constructed from
file metadata until the corresponding validator has passed.}

Implement the pattern by component:

\begin{minted}{c}
herr_t H5O_msg_decode_raw(const uint8_t *buf, size_t buf_size, H5O_msg_raw_t *raw);
herr_t H5O_msg_validate(const H5F_t *f, const H5O_msg_raw_t *raw, H5O_validation_ctx_t *ctx);
herr_t H5O_msg_construct(const H5O_msg_raw_t *raw, H5O_msg_t *out);    
\end{minted}

Do not let decode functions allocate complex structures before basic
bounds are known. A raw decoded object should hold values, offsets,
and spans, not trusted pointers into long-lived internal state.

All decode functions that consume file bytes should take one of these forms:
\begin{minted}{c}
herr_t decode(const H5F_t *f, const uint8_t *buf, size_t buf_size, out_t *out);

herr_t decode_cursor(H5_decode_cursor_t *cur, out_t *out);
\end{minted}

Ban this pattern in security-sensitive decode code:
\begin{minted}{c}
const uint8_t **pp
\end{minted}
unless there is also an explicit \texttt{p\_end} or remaining-size counter. The cache-image fix already moved in this direction
by adding a buffer-size argument and bounds checks.~\cite{commit_3914bb7}

\subsubsection*{Priority 3: Centralize checked arithmetic and allocation}

Most parser bugs become exploitable when metadata reaches allocation, copy, or pointer arithmetic. Add a small, mandatory checked-arithmetic layer.

Example API:
\begin{minted}{c}
herr_t H5_checked_add_hsize(hsize_t a, hsize_t b, hsize_t *out);
herr_t H5_checked_mul_hsize(hsize_t a, hsize_t b, hsize_t *out);
herr_t H5_checked_addr_add(haddr_t base, hsize_t len, haddr_t eoa, haddr_t *end);
herr_t H5_checked_alloc_size(size_t count, size_t elem_size, size_t *bytes);
\end{minted}

Then make these rules review-blocking:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.48\linewidth}p{0.48\linewidth}@{}}
\caption{Review-blocking parser arithmetic and allocation patterns}
\label{tab:review-blocking-parser-patterns}\\
\hline
\textbf{Ban} & \textbf{Replacement} \\
\hline
\endfirsthead

\caption[]{Review-blocking parser arithmetic and allocation patterns, continued}\\
\hline
\textbf{Ban} & \textbf{Replacement} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

\code{malloc(n * size)} & \code{H5MM_calloc_checked(n, size)} \\
\code{addr + size} & \code{H5_checked_addr_add(addr, size, eoa, \&end)} \\
\code{memcpy(dst, src, n)} from file-derived size &
\code{H5_checked_memcpy(dst, dst_size, src, src_size, n)} \\
pointer increment without \code{p_end} & cursor-based decode \\
raw recursion through object graph & bounded traversal context \\
\end{longtable}
\endgroup

The 2024 security commit added multiplication-overflow checks for
dataset storage size and checks against file bounds.~\cite{commit_ce53bc0}
That pattern should become a library-wide rule, not a local repair.

\subsubsection*{Priority 4: Create security profiles for file opening}

Add explicit modes. Do not rely on users to infer safe behavior.
\begin{minted}{c}
typedef enum {
    H5_SECURITY_PROFILE_LEGACY,
    H5_SECURITY_PROFILE_TRUSTED_FAST,
    H5_SECURITY_PROFILE_UNTRUSTED_STRICT,
    H5_SECURITY_PROFILE_FORENSIC
} H5_security_profile_t;
\end{minted}

Suggested semantics:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.24\linewidth}p{0.72\linewidth}@{}}
\caption{Security profile semantics}
\label{tab:security-profile-semantics}\\
\hline
\textbf{Profile} & \textbf{Behavior} \\
\hline
\endfirsthead

\caption[]{Security profile semantics, continued}\\
\hline
\textbf{Profile} & \textbf{Behavior} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

\code{legacy} & current compatibility bias \\
\code{trusted_fast} & normal validation, generous resource limits \\
\code{untrusted_strict} & strict metadata validation, plugin loading disabled unless allowlisted, resource limits enforced \\
\code{forensic} & never repairs silently, reports all structural anomalies, best-effort inspection without unsafe traversal \\
\end{longtable}
\endgroup

Expose it through file-access property lists:
\begin{minted}{c}
H5Pset_security_profile(fapl, H5_SECURITY_PROFILE_UNTRUSTED_STRICT);
H5Pset_max_metadata_bytes(fapl, 256 * 1024 * 1024);
H5Pset_max_object_header_chunks(fapl, 1024);
H5Pset_max_btree_depth(fapl, 64);
H5Pset_max_filter_expansion_ratio(fapl, 100);
H5Pset_plugin_policy(fapl, H5_PLUGIN_POLICY_DISABLED);
\end{minted}

This matters because HDF5 supports dynamically loaded plugins, and the
runtime API can enable or disable plugin classes; the docs also state
that \texttt{HDF5\_PLUGIN\_PRELOAD=::} disables all dynamic plugins.
Turn that capability into a first-class security policy, not an
environment-variable trick.

\subsubsection*{Priority 5: Fuzz by structure, not just by whole-file crashes}

Whole-file fuzzing is necessary but insufficient. It finds symptoms. Structure-aware fuzzing finds families.

Build fuzz targets in this order:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}r p{0.40\linewidth}p{0.48\linewidth}@{}}
\caption{Structure-aware fuzzing priorities}
\label{tab:structure-aware-fuzzing-priorities}\\
\hline
\textbf{Order} & \textbf{Fuzz target} & \textbf{Why first} \\
\hline
\endfirsthead

\caption[]{Structure-aware fuzzing priorities, continued}\\
\hline
\textbf{Order} & \textbf{Fuzz target} & \textbf{Why first} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

1 & object-header message decoder & many paths flow through it \\
2 & datatype decoder and converter & high exploitability and complex nested metadata \\
3 & dataspace and selection decoder & rank, count, hyperslab, and point math \\
4 & layout and chunk index decoder & storage-size, address, and chunk-product risks \\
5 & free-space and local/global heap decoders & repeated historical overflow and list-corruption class \\
6 & filter pipeline and scale-offset filter & compression/decompression boundary risk \\
7 & metadata cache image reconstruction & high complexity and recent CVE pattern \\
8 & \code{h5dump}/\code{h5repack} rendering paths & tool paths often bypass library assumptions \\
\end{longtable}
\endgroup

Use three fuzzing layers:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.28\linewidth}p{0.68\linewidth}@{}}
\caption{Fuzzing layers}
\label{tab:fuzzing-layers}\\
\hline
\textbf{Layer} & \textbf{Description} \\
\hline
\endfirsthead

\caption[]{Fuzzing layers, continued}\\
\hline
\textbf{Layer} & \textbf{Description} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

byte-level fuzzing & current randomized mutation against APIs \\
structure-aware mutation & valid superblock with corrupted object headers, valid object header with bad continuation, valid datatype with bad member offset \\
invariant-guided generation & generate files that violate exactly one invariant at a time \\
\end{longtable}
\endgroup

OSS-Fuzz already supports libFuzzer, AFL++, Honggfuzz, Centipede, sanitizers,
and distributed execution for C/C++ projects. Use OSS-Fuzz for long-running
public coverage and ClusterFuzzLite or local CI for PR-level checks.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.24\linewidth}p{0.72\linewidth}@{}}
\caption{Fuzzing acceptance criteria}
\label{tab:fuzzing-acceptance-criteria}\\
\hline
\textbf{Stage} & \textbf{Gate} \\
\hline
\endfirsthead

\caption[]{Fuzzing acceptance criteria, continued}\\
\hline
\textbf{Stage} & \textbf{Gate} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

PR & run minimized CVE corpus under \code{ASan}/\code{UBSan}/\code{LSan} \\
nightly & run structure fuzzers for core parsers \\
weekly & report coverage by component, not just total line coverage \\
release candidate & 72-hour fuzz soak on changed parser areas \\
CVE closure & crashing input added to corpus and mapped to invariant \\
\end{longtable}
\endgroup

\subsubsection*{Priority 6: Make malformed-input regression cheap}

Create this repository layout:
\begin{verbatim}
test/security/
  corpus/
    cve/
      CVE-2025-7069/
        input.h5
        expected.yml
      CVE-2025-6269/
        input.h5
        expected.yml
    invariants/
      H5O-CONT-001/
      H5T-SIZE-002/
  harnesses/
    fuzz_object_header.c
    fuzz_datatype.c
    fuzz_dataspace.c
    fuzz_layout.c
    fuzz_filters.c
  expected_errors/
    corrupt_file.yml
\end{verbatim}

Each expected.yml should include:
\begin{minted}{yaml}
expected:
  exit: controlled_error
  allowed_errors:
    - H5E_CORRUPT
    - H5E_BADVALUE
    - H5E_OVERFLOW
forbidden:
  - SIGSEGV
  - SIGABRT
  - leak
  - use_after_free
  - timeout
invariant:
  - H5FS-PAGE-001
\end{minted}

This gives maintainers a fast way to prove that old crashes remain dead.

\subsubsection*{Priority 7: Add a security-review checklist that blocks risky parser changes}

Every PR touching decode, encode, object lifecycle, filters, tools, or plugin
loading should answer:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.58\linewidth}p{0.38\linewidth}@{}}
\caption{Security-review checklist for parser-adjacent changes}
\label{tab:security-review-checklist}\\
\hline
\textbf{Question} & \textbf{Required answer} \\
\hline
\endfirsthead

\caption[]{Security-review checklist for parser-adjacent changes, continued}\\
\hline
\textbf{Question} & \textbf{Required answer} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

Does this read file-derived bytes? & list buffer and bounds source \\
Does this allocate using file-derived values? & show checked multiplication \\
Does this copy memory using file-derived sizes? & show source and destination bounds \\
Does this traverse links, chunks, heaps, B-trees, or continuations? & show cycle and depth bounds \\
Does this construct native objects before validation? & explain why or refactor \\
Does this add or change cleanup paths? & show partial-init safety \\
Does this invoke plugins or filters? & show policy interaction \\
Does this alter error handling? & show malformed-input test \\
\end{longtable}
\endgroup

Add automated enforcement where possible:

\begin{verbatim}
fail PR if new decode function lacks size parameter
fail PR if malloc(count * size) appears outside checked wrapper
fail PR if memcpy appears in parser code without checked wrapper annotation
fail PR if new parser file lacks fuzz target registration
fail PR if new object traversal lacks max-depth or visited-set logic
\end{verbatim}

This will annoy contributors at first. Good. That friction is cheaper than post-release CVEs.

\subsubsection*{Priority 8: Separate compatibility from safety}

HDF5’s backward compatibility is valuable, but it creates a security trap: old,
permissive parsing keeps malformed legacy structures alive.

Implement this policy:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.42\linewidth}p{0.54\linewidth}@{}}
\caption{Compatibility and safety profile policy}
\label{tab:compatibility-safety-profile-policy}\\
\hline
\textbf{Case} & \textbf{Behavior} \\
\hline
\endfirsthead

\caption[]{Compatibility and safety profile policy, continued}\\
\hline
\textbf{Case} & \textbf{Behavior} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

valid legacy file & accepted \\
ambiguous legacy construct & accepted only in \code{legacy} or \code{trusted_fast}; warning in \code{forensic}; rejected in \code{untrusted_strict} \\
structurally invalid file that old versions tolerated & rejected in all non-legacy profiles \\
parser repair needed & repair only through explicit tool, never silently during normal open \\
unknown message with unsafe flags & rejected unless profile explicitly allows preservation \\
\end{longtable}
\endgroup

The point is not to break legitimate archives. The point is to stop treating
corrupted metadata as compatibility.

\subsubsection*{Priority 9: Harden tools as separate attack surfaces}

Do not assume \code{h5dump}, \code{h5repack}, converters, and test utilities are
less important. They are often the first thing users run on untrusted files.

Actions:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.30\linewidth}p{0.66\linewidth}@{}}
\caption{Tool hardening actions}
\label{tab:tool-hardening-actions}\\
\hline
\textbf{Tool class} & \textbf{Change} \\
\hline
\endfirsthead

\caption[]{Tool hardening actions, continued}\\
\hline
\textbf{Tool class} & \textbf{Change} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

inspection tools & use \code{H5_SECURITY_PROFILE_FORENSIC} by default \\
conversion tools & use \code{H5_SECURITY_PROFILE_UNTRUSTED_STRICT} by default \\
rendering paths & bounded strings only \\
legacy converters & disable by default; build behind explicit \code{HDF5_ENABLE_LEGACY_UNSAFE_TOOLS=ON} \\
plugin-dependent tools & warn or require allowlist \\
\end{longtable}
\endgroup

A dangerous file should not become more dangerous because the user tried to
inspect it.

\subsubsection*{Priority 10: Plugin policy and signed extensions}

Plugins are a different risk class. They are executable code loaded at runtime.
HDF5’s plugin system supports dynamically loaded filter plugins and lets users
implement custom filters. That is useful, but it must not be invisible in
untrusted mode.

Implement:
\begin{verbatim}
default profile:
  trusted_fast: existing plugin behavior
  untrusted_strict: plugins disabled unless allowlisted
  forensic: plugins disabled; report missing filter
\end{verbatim}

Add a plugin manifest:
\begin{minted}{yaml}
plugin_id: 32008
name: blosc
version: 1.0.0
sha256: ...
signer: ...
allowed_profiles:
  - trusted_fast
  - untrusted_strict
capabilities:
  - filter_decode
limits:
  max_expansion_ratio: 100
  max_alloc_bytes: 1073741824
\end{minted}

Do not start with a perfect PKI. Start with allowlisted hashes and paths.
Then add signatures. The signed-plugin work matters, but without resource limits
a signed buggy plugin can still crash the process.

\subsubsection*{Priority 11: Supply-chain controls after parser controls}

Supply-chain work matters, but it is second-order relative to malformed-file parsing. Do it in parallel only after the parser program has owners.

Minimum controls:

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.32\linewidth}p{0.64\linewidth}@{}}
\caption{Minimum supply-chain controls}
\label{tab:minimum-supply-chain-controls}\\
\hline
\textbf{Control} & \textbf{Target} \\
\hline
\endfirsthead

\caption[]{Minimum supply-chain controls, continued}\\
\hline
\textbf{Control} & \textbf{Target} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

OpenSSF Scorecard & weekly scheduled run, alerts in security tab \\
CodeQL & required for PRs touching C/C++ parser surfaces \\
pinned GitHub Actions & no floating third-party action refs \\
SBOM & release artifacts include source and binary SBOMs \\
signed release artifacts & signatures plus checksums \\
SLSA provenance & release builds produce verifiable provenance \\
\end{longtable}
\endgroup

OpenSSF Scorecard has an official GitHub Action, and public repositories can use
it without cost. The SLSA GitHub generator provides tooling for SLSA provenance
on GitHub Actions, including provenance that users can verify to understand how
an artifact was built.

\subsubsection*{Priority 12: Metrics that force the right behavior}

Use metrics that measure class elimination, not activity.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\begin{longtable}{@{}p{0.70\linewidth}p{0.26\linewidth}@{}}
\caption{Security program metrics}
\label{tab:security-program-metrics}\\
\hline
\textbf{Metric} & \textbf{Target} \\
\hline
\endfirsthead

\caption[]{Security program metrics, continued}\\
\hline
\textbf{Metric} & \textbf{Target} \\
\hline
\endhead

\hline
\endfoot

\hline
\endlastfoot

CVE corpus pass rate under ASan/UBSan/LSan & 100\% \\
parser decode functions with explicit buffer size or cursor & 100\% \\
parser allocations through checked wrappers & 100\% \\
parser memcpy/memmove through checked wrappers or reviewed exemption & 100\% \\
high-risk structures with invariant files & 100\% \\
high-risk structures with fuzz target & 100\% \\
new CVEs that add invariant updates & 100\% \\
time from new crash to minimized corpus seed & under 48 hours \\
time from minimized seed to invariant classification & under 5 business days \\
sanitizer-clean nightly fuzz runs & sustained \\
untrusted profile test coverage & every release \\
\end{longtable}
\endgroup

Track this in a public dashboard. Not because dashboards fix bugs. They stop the
team from pretending that “more tests” is a plan.

\subsubsection{Privacy}

In this section we summarize privacy threats in HDF5 based on the audit findings and 
describe mitigation strategies and their priorities.

\paragraph{Overview of Privacy Threat Categories:}

The HDF5 Privacy Exposure Model ~\cite{PVEM} identifies 
seven categories of privacy threats that describe the principal mechanisms through which 
sensitive information may be unintentionally exposed in HDF5. Privacy threats primarily 
arise during normal scientific data creation, processing, sharing, and preservation. 
They are a consequence of HDF5's self-describing data model, rich metadata capabilities, 
extensible software architecture, broad interoperability across scientific applications, 
and prolonged data life cycle.

Several threat categories involve direct disclosure of information stored within HDF5 files. 
Metadata Exposure (P1) occurs when descriptive information, such as object names, attributes, 
datatype member names, or other user-defined metadata, reveals sensitive contextual 
information even when the underlying scientific data are protected. Structural Metadata 
Exposure (P2) arises from examination of HDF5's internal storage structures, 
including file layout, storage organization, and implementation metadata, which may 
disclose characteristics of the stored data without requiring HDF5 software. 
Raw Data Exposure (P3) addresses situations in which information about scientific data 
becomes observable through storage patterns, embedded textual descriptions, or 
other implementation artifacts rather than through intended data access.

Another group of threats results from routine scientific data processing. Unintended 
Data Alteration or Inclusion (P4) occurs when software automatically generates, expands, 
or preserves additional information during data conversion or processing. 
Data Aggregation and Correlation (P5) recognizes that individually non-sensitive 
datasets may collectively reveal sensitive information when combined or interpreted 
together. Persistent Data Beyond Intended Usage (P6) addresses privacy risks 
associated with temporary files, metadata snapshots, cached datasets, 
archived copies, and other artifacts that remain accessible after their 
intended purpose has ended.

The remaining threat category Misplaced Trust in Privacy by format (P7) reflects broader 
characteristics of modern HDF5 ecosystems. Examples include: 

\begin{itemize}

\item Misinterpretation of data privacy in HDF5 format leads to the possibility of 
sensitive information disclosure through metadata, structural information, or inference.

\item Conversion to HDF5 highlights that transforming legacy scientific formats 
into HDF5 often increases accessibility and interoperability by making implicit 
information explicit, but this additional semantic information may itself introduce new privacy risks. 

\item Loss of Data Control describes situations in which data movement, replication, 
backup, cloud migration, or transfer across organizational boundaries weakens the 
original privacy protections and governance mechanisms applied to the data.

\end{itemize}

Although presented as distinct categories, these threats frequently occur together. 
A single scientific workflow may enrich metadata during format conversion, aggregate 
datasets from multiple sources, generate temporary processing artifacts, and 
replicate results across multiple storage systems. Consequently, privacy risks should 
be evaluated across the complete scientific data lifecycle rather than 
by considering individual HDF5 files in isolation.

\paragraph{Mitigation Priorities.}
Because HDF5 was designed to maximize scientific data sharing, portability, 
interoperability, and long-term usability rather than enforce privacy, 
no single mechanism can eliminate all privacy risks. Effective privacy protection 
therefore requires a multi-facet approach that considers the entire scientific data 
lifecycle, including data creation, processing, storage, sharing, transfer, 
and archival. Privacy should be treated as a property of the complete HDF5 ecosystem 
including applications, plugins, workflows, storage systems, and organizational 
practices rather than solely of the HDF5 file format. 
Here we list mitigation strategies and their priorities.

\paragraph{Priority 1: \emph{Privacy exposure awareness.}}
Scientists, software developers, data stewards, and infrastructure providers should 
recognize that HDF5's self-describing nature inherently exposes contextual 
information in many ways. Privacy reviews should therefore evaluate not only 
scientific datasets, but also metadata, structural organization, workflow artifacts, 
derived products, and information that may be inferred through aggregation or 
correlation. Reviews should also take into consideration possible privacy 
exposure during data lifecycle. 

\paragraph{Priority 2: \emph{Privacy exposure through HDF5 tools and plugins awareness.}}
The HDF5 ecosystem extends beyond the core HDF5 software (i.e., file format and the library), 
and includes a diverse collection of command-line utilities, visualization tools, 
analysis libraries, VOL connectors, VFDs, filters, and third-party applications. 
These components frequently generate additional outputs such as logs, diagnostic 
messages, temporary files, metadata snapshots, converted datasets, and cached 
intermediate results that may expose sensitive information beyond the contents of the 
original HDF5 file. Furthermore, plugins may introduce new metadata, or transform 
data in ways that alter their privacy characteristics. Consequently, privacy 
assessments should evaluate not only HDF5 files but also the behavior of the 
software ecosystem that processes them. Developers should document the privacy 
implications of their tools, minimize unnecessary generation of sensitive artifacts, 
and provide users with mechanisms to control logging, metadata generation, and 
retention of intermediate products.

\paragraph{Priority 3: \emph{Privacy-aware system design.}}
HDF5 software systems should be built with data privacy in mind. Applications should 
minimize collection and retention of unnecessary metadata, avoid embedding sensitive 
information in object names or descriptive attributes, remove temporary processing 
artifacts when they are no longer needed, and carefully evaluate automatically 
generated metadata introduced during data conversion or interoperability. 
Workflow designers should review intermediate products, logs, cached datasets, and 
derived files before publication or long-term archival. 

\paragraph{Priority 4: \emph{Data encryption as privacy control.}}
Encryption represents an important but limited privacy control. 

Currently, HDF5 doesn’t support data encryption that allows I/O on encrypted 
data, effectively disabling sensitive data exposure during data 
transfer and storage. Users may implement custom raw data encryption 
and partial metadata encryption (i.e., names of the objects) using 
existing HDF5 filtering mechanism. Such approach still leaves user-defined 
metadata (e.g., attributes) and structural metadata unencrypted. Information 
that remains unencrypted as application-specific workflow artifacts, 
temporary files, system logs, or derived datasets may continue to 
disclose sensitive information.
  
HDF5 encryption feature that properly implements encryption of all data 
stored in HDF5 while allowing partial I/O on encrypted data is currently 
under development. However, encryption does not eliminate all HDF5 privacy risks. 
Once encrypted data are decrypted for legitimate processing, applications, 
plugins, caches, diagnostic logs, and intermediate workflow products may again 
become sources of unintended exposure. Encryption therefore reduces the attack 
surface but does not address inference, aggregation, excessive metadata 
generation, data retention, or loss of governance. Encryption also complicates 
workflows and may lead to data loss if on some stage of data life cycle 
encryption keys become unavailable.

Therefore, encryption should be viewed as one component of a comprehensive 
privacy strategy rather than as a complete solution. Effective protection 
requires combining encryption with careful metadata management, lifecycle-aware 
data governance, privacy-aware application design, access controls provided by 
the underlying computing environment, and education of users regarding the 
unique privacy characteristics of HDF5-based scientific data systems.


\end{document}
