\documentclass[../main.tex]{subfiles}

\begin{document}

%% ======================================================================
%%  Section 7 – HDF5 in the Software Supply Chain
%% ======================================================================
\section{HDF5 in the Software Supply Chain}
\label{sec:supply-chain}

\begin{itemize}
    \item HDF5 2.0.0 source-artifact integrity check~\cite{prisupp-2204}
    \item SBOM
\end{itemize}

\subsection{Upstream Dependencies}

\providecommand{\depcode}[1]{\nolinkurl{#1}}

This subsection summarizes external software that can be required by an HDF5
build. It is organized by build option, because the default serial C library has
a much smaller dependency set than feature-complete, testing, packaging, or
developer builds.

\subsubsection{Dependency Source Modes}

HDF5's CMake build can either locate dependencies that are already installed or,
for selected filter-related projects, add external source trees to the build with
\depcode{FetchContent}.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.19\linewidth}p{0.34\linewidth}p{0.38\linewidth}@{}}
\hline
\textbf{Mode} & \textbf{Relevant options} & \textbf{Meaning} \\
\hline
System packages & \depcode{HDF5_ALLOW_EXTERNAL_SUPPORT=NO} &
CMake uses \depcode{find_package()}, \depcode{find_library()}, or
\depcode{find_program()} to locate installed dependencies. This is the default. \\
Git fetch & \depcode{HDF5_ALLOW_EXTERNAL_SUPPORT=GIT} plus a
dependency-specific \depcode{*_USE_EXTERNAL=ON} option &
CMake downloads source from the configured \depcode{*_GIT_URL} and
\depcode{*_GIT_TAG} values and builds it with HDF5. \\
Tarball fetch & \depcode{HDF5_ALLOW_EXTERNAL_SUPPORT=TGZ} plus a
dependency-specific \depcode{*_USE_EXTERNAL=ON} option &
CMake downloads from \depcode{*_TGZ_ORIGPATH} and \depcode{*_TGZ_NAME}, or uses
\depcode{TGZPATH} when the dependency-specific \depcode{*_USE_LOCALCONTENT=ON}
option is set. \\
\hline
\end{longtable}
\endgroup

The \depcode{HDF5_ALLOW_EXTERNAL_SUPPORT} modes apply to filter-related
dependencies. Some test and example configurations have their own source-fetch
paths, such as KWSys for the API test driver and h5py for Python examples. MPI,
HDFS, ROS3, OpenSSL, Java, documentation, packaging, and developer tools must be
provided by the build environment.

\subsubsection{Baseline Build}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.24\linewidth}p{0.22\linewidth}p{0.45\linewidth}@{}}
\hline
\textbf{Dependency} & \textbf{Trigger} & \textbf{Notes} \\
\hline
CMake 3.26 or newer & Always &
Required by the top-level \depcode{cmake_minimum_required()}. \\
C compiler and native build tool & Always &
The top-level project is a C project. \\
Platform C library, headers, and system calls & Always &
CMake probes standard headers, types, functions, byte order, large-file support,
file locking, \depcode{pread}/\depcode{pwrite}, and related platform features. \\
Platform link libraries & Platform-dependent &
CMake adds libraries only when probes show they are needed: \depcode{m},
\depcode{dl}, \depcode{rt}, \depcode{posix4}, \depcode{ucb}, \depcode{ws2_32},
\depcode{wsock32}, and \depcode{shlwapi}. \\
PkgConfig & Optional probe &
Used only when available. It can help some find modules, such as zlib-ng and
libaec, but it is not a hard requirement for the default build. \\
Threads & Found in every configure; required by thread features &
CMake searches for C11 threads, Win32 threads, or pthreads. When found, HDF5
records thread support and links \depcode{Threads::Threads};
\depcode{HDF5_ENABLE_THREADSAFE}, \depcode{HDF5_ENABLE_CONCURRENCY}, and
\depcode{HDF5_ENABLE_SUBFILING_VFD} require it. \\
\hline
\end{longtable}
\endgroup

\subsubsection{Feature Dependencies}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.20\linewidth}p{0.27\linewidth}p{0.20\linewidth}p{0.24\linewidth}@{}}
\hline
\textbf{Feature or component} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Parallel HDF5 & \depcode{HDF5_ENABLE_PARALLEL=ON} &
MPI C with MPI-IO, MPI standard 3.0 or newer &
Adds \depcode{MPI::MPI_C} to public link libraries. Parallel filtered writes and
large parallel I/O are enabled only when the required MPI symbols are present.
Parallel tests also need an MPI launcher such as \depcode{mpiexec}. \\
Parallel Fortran & \depcode{HDF5_ENABLE_PARALLEL=ON},
\depcode{HDF5_BUILD_FORTRAN=ON} &
MPI Fortran &
The Fortran build adds MPI Fortran libraries. The \depcode{mpi_f08} module is
used when available. \\
Thread-safe library & \depcode{HDF5_ENABLE_THREADSAFE=ON} &
Threads &
Unsupported with parallel, Fortran, C++, or high-level library unless
\depcode{HDF5_ALLOW_UNSUPPORTED=ON}. On Windows, static libraries are rejected. \\
Multi-threaded concurrency & \depcode{HDF5_ENABLE_CONCURRENCY=ON} &
Threads &
Experimental. Mutually exclusive with \depcode{HDF5_ENABLE_THREADSAFE}; has the
same unsupported combinations as thread-safety unless overridden with
\depcode{HDF5_ALLOW_UNSUPPORTED=ON}. \\
C++ wrappers & \depcode{HDF5_BUILD_CPP_LIB=ON} &
C++ compiler &
Unsupported with parallel unless \depcode{HDF5_ALLOW_UNSUPPORTED=ON}.
High-level C++ wrappers also require \depcode{HDF5_BUILD_HL_LIB=ON}. \\
Fortran wrappers & \depcode{HDF5_BUILD_FORTRAN=ON} &
Fortran compiler with Fortran 2003 \depcode{ISO_C_BINDING} support &
Unsupported with thread-safety or concurrency unless
\depcode{HDF5_ALLOW_UNSUPPORTED=ON}. High-level Fortran wrappers also require
\depcode{HDF5_BUILD_HL_LIB=ON}. \\
Java wrappers & \depcode{HDF5_BUILD_JAVA=ON} &
Java/JDK, shared HDF5 libraries, Java dependency JARs &
Java requires shared HDF5 libraries. Java 11 or newer is required for the JNI
implementation. Java tests additionally require the JUnit and Hamcrest JARs. \\
Java JNI implementation & \depcode{HDF5_BUILD_JAVA=ON},
\depcode{HDF5_ENABLE_JNI=ON}, or any Java version earlier than 25 &
JNI headers and libraries &
JNI is the default Java implementation. HDFS also requires JNI. \\
Java FFM implementation & \depcode{HDF5_BUILD_JAVA=ON}, Java 25 or newer,
\depcode{HDF5_ENABLE_JNI=OFF} &
\depcode{jextract} from \depcode{JEXTRACT_HOME/bin} or \depcode{JAVA_HOME/bin} &
Generates FFM bindings at configure time. \\
Java logging/test JARs & \depcode{HDF5_BUILD_JAVA=ON} &
\depcode{slf4j-api}, \depcode{slf4j-nop}, \depcode{slf4j-simple}; plus JUnit 4
and Hamcrest for tests &
The default paths point to bundled JARs under \depcode{java/lib}. They can be
overridden with \depcode{HDF5_JAVA_*_JAR} cache variables.
\depcode{HDF5_ENABLE_MAVEN_DEPLOY=ON} generates Maven-style artifacts and POM
files; the CMake build does not locate or run \depcode{mvn}. \\
Documentation & \depcode{HDF5_BUILD_DOC=ON} &
Doxygen &
\depcode{HDF5_ENABLE_DOXY_WARNINGS=ON} makes Doxygen warnings fail the build. \\
Header regeneration & \depcode{HDF5_GENERATE_HEADERS=ON} &
Perl &
Perl is optional otherwise. If it is missing, generated headers are not
regenerated. \\
Python examples & \depcode{H5EXAMPLE_BUILD_PYTHON=ON} &
Python3 interpreter, Python development files, NumPy, h5py source, pip &
The examples CMake project fetches \depcode{h5py} from Git and installs it with
\texttt{python3 -m pip}. \\
\hline
\end{longtable}
\endgroup

\subsubsection{Filter Dependencies}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.19\linewidth}p{0.28\linewidth}p{0.22\linewidth}p{0.22\linewidth}@{}}
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Source choices} \\
\hline
DEFLATE filter & \depcode{HDF5_ENABLE_ZLIB_SUPPORT=ON},
\depcode{HDF5_USE_ZLIB_NG=OFF} &
zlib &
System package by default; \depcode{ZLIB_USE_EXTERNAL=ON} can build zlib from
Git or TGZ. \depcode{HDF5_USE_ZLIB_STATIC=ON} prefers a static library. \\
DEFLATE filter through zlib-ng & \depcode{HDF5_ENABLE_ZLIB_SUPPORT=ON},
\depcode{HDF5_USE_ZLIB_NG=ON} &
zlib-ng &
System package by default; \depcode{ZLIB_USE_EXTERNAL=ON} can build zlib-ng from
Git or TGZ. zlib compatibility mode is detected. \\
SZIP filter & \depcode{HDF5_ENABLE_SZIP_SUPPORT=ON} &
libaec with libsz compatibility, or another package selected by
\depcode{LIBAEC_PACKAGE_NAME} &
System package by default; \depcode{SZIP_USE_EXTERNAL=ON} can build libaec from
Git or TGZ. \depcode{HDF5_USE_LIBAEC_STATIC=ON} prefers static libraries. \\
HDF5 filter plugins project & \depcode{HDF5_ENABLE_PLUGIN_SUPPORT=ON} &
Installed \depcode{h5pl}/HDF5 filter plugins package, or the
\depcode{hdf5_plugins} project &
\depcode{PLUGIN_USE_EXTERNAL=ON} can build the plugin project from Git or TGZ.
Plugin support is disabled for the 1.6 API. \\
\hline
\end{longtable}
\endgroup

The plugin project can introduce additional filter-library dependencies. The
HDF5 cache defaults include entries for bitgroom, bitround, bitshuffle, Blosc,
Blosc2, bzip2, fpzip, JPEG, LZ4, LZF, SZ, ZFP, and Zstandard. Those libraries
are controlled by the plugin project's own options when \depcode{hdf5_plugins}
is built or found.

\subsubsection{Virtual File Drivers and External Connectors}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.19\linewidth}p{0.28\linewidth}p{0.22\linewidth}p{0.22\linewidth}@{}}
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Direct VFD & \depcode{HDF5_ENABLE_DIRECT_VFD=ON} &
POSIX \depcode{O_DIRECT} and \depcode{posix_memalign()} support &
No third-party library. Configuration fails if the platform support is missing. \\
Mirror VFD & \depcode{HDF5_ENABLE_MIRROR_VFD=ON} &
POSIX socket/fork headers and functions &
No third-party library. Requires \depcode{netinet/in.h}, \depcode{netdb.h},
\depcode{arpa/inet.h}, \depcode{sys/socket.h}, and \depcode{fork()}. \\
ROS3 VFD & \depcode{HDF5_ENABLE_ROS3_VFD=ON} &
\depcode{aws-c-s3} CMake package &
HDF5 does not fetch this library. Building \depcode{aws-c-s3} from source can
also require AWS C libraries such as \depcode{aws-c-common},
\depcode{aws-checksums}, \depcode{aws-c-cal}, \depcode{aws-c-io},
\depcode{aws-c-compression}, \depcode{aws-c-http}, \depcode{aws-c-sdkutils},
\depcode{aws-c-auth}, and, on Linux, \depcode{aws-lc} and \depcode{s2n-tls}.
ROS3 docker-proxy tests additionally require \depcode{docker} and the AWS CLI. \\
HDFS VFD & \depcode{HDF5_ENABLE_HDFS=ON} &
JNI/JVM and Hadoop \depcode{libhdfs} &
\depcode{HADOOP_HOME} is used to locate \depcode{hdfs.h}, \depcode{libhdfs},
and the Hadoop version command. Non-MSVC builds also add \depcode{-pthread}. \\
Subfiling VFD and IOC VFD & \depcode{HDF5_ENABLE_PARALLEL=ON},
\depcode{HDF5_ENABLE_SUBFILING_VFD=ON} &
MPI C with \depcode{MPI_Comm_split_type}, plus Threads &
Not available on Windows. IOC VFD is built when subfiling is enabled. \\
External VOL connectors & \depcode{HDF5_VOL_ALLOW_EXTERNAL=GIT} or
\depcode{LOCAL_DIR} with \depcode{HDF5_VOL_URLnn} or \depcode{HDF5_VOL_PATHnn} &
Connector-defined &
Up to ten CMake-based VOL connectors can be added. HDF5 patches connector CMake
code to use the in-tree HDF5 build, but each connector may bring its own
dependencies. \\
\hline
\end{longtable}
\endgroup

\subsubsection{Tools, Tests, and Security Features}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.21\linewidth}p{0.29\linewidth}p{0.20\linewidth}p{0.21\linewidth}@{}}
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Signed plugin verification & \depcode{HDF5_REQUIRE_SIGNED_PLUGINS=ON} &
OpenSSL libcrypto 1.1.0 or newer &
Adds \depcode{OpenSSL::Crypto} to HDF5. \depcode{HDF5_LOCK_PLUGIN_KEYSTORE=ON}
also requires \depcode{HDF5_PLUGIN_KEYSTORE_DIR}. \\
\depcode{h5sign} tool & \depcode{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\depcode{HDF5_BUILD_TOOLS=ON} &
OpenSSL libcrypto &
Used to sign HDF5 plugins. \\
Signed-plugin tests & \depcode{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\depcode{BUILD_TESTING=ON} &
\depcode{openssl} executable &
Tests generate RSA keys with the OpenSSL command-line tool. \\
Parallel tools utilities & \depcode{HDF5_ENABLE_PARALLEL=ON},
\depcode{HDF5_BUILD_PARALLEL_TOOLS=ON} &
MFU, CIRCLE, DTCMP &
CMake looks for headers and libraries, using \depcode{MFU_ROOT} as a hint. \\
API test driver & \depcode{BUILD_TESTING=ON},
\depcode{HDF5_TEST_API_ENABLE_DRIVER=ON} &
KWSys &
The test driver fetches KWSys from a tarball or from \depcode{TGZPATH} when
\depcode{KWSYS_USE_LOCALCONTENT=ON}. \\
Shell-script tests & \depcode{BUILD_TESTING=ON} &
PowerShell or Bash &
CMake uses \depcode{pwsh}/\depcode{powershell} when available; otherwise Unix
builds look for \depcode{bash}. \\
ROS3 VFD tests & \depcode{HDF5_ENABLE_ROS3_VFD=ON},
\depcode{HDF5_ENABLE_ROS3_VFD_DOCKER_PROXY=ON} &
Docker daemon and AWS CLI &
Used to run S3 proxy tests. \\
\hline
\end{longtable}
\endgroup

\subsubsection{Developer and Packaging Dependencies}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.25\linewidth}p{0.34\linewidth}p{0.32\linewidth}@{}}
\hline
\textbf{Configuration} & \textbf{Build options} & \textbf{Dependency} \\
\hline
Analyzer tooling & \depcode{HDF5_ENABLE_ANALYZER_TOOLS=ON} &
\depcode{clang-tidy}, \depcode{include-what-you-use}, and \depcode{cppcheck}
are probed and used by the analyzer macros. \\
Source formatting targets & \depcode{HDF5_ENABLE_FORMATTERS=ON} &
\depcode{clang-format} and \depcode{cmake-format}. \\
Code coverage with Clang/IntelLLVM & \depcode{HDF5_ENABLE_COVERAGE=ON},
\depcode{CODE_COVERAGE=ON} &
\depcode{llvm-cov} and \depcode{llvm-profdata}; AppleClang uses
\texttt{xcrun llvm-cov} and \texttt{xcrun llvm-profdata}. \\
Code coverage with GCC & \depcode{HDF5_ENABLE_COVERAGE=ON},
\depcode{CODE_COVERAGE=ON} &
\depcode{lcov} and \depcode{genhtml}; fallback instrumentation can also link
\depcode{gcov}. \\
Sanitizers & \depcode{HDF5_ENABLE_SANITIZERS=ON},
\depcode{HDF5_USE_SANITIZER=<kind>} &
Compiler sanitizer runtime support. Supported values depend on compiler and
platform. \\
AFL support module & Explicit inclusion of
\depcode{config/sanitizer/afl-fuzzing.cmake} &
\depcode{afl-cc} and \depcode{afl-c++}. This module is available but is not
included by the top-level HDF5 build by default. \\
Dependency graph support module & Explicit inclusion of
\depcode{config/sanitizer/dependency-graph.cmake} &
Graphviz \depcode{dot}. This module is available but is not included by the
top-level HDF5 build by default. \\
Windows installers & CPack on Windows &
NSIS and WiX are probed when available. \\
Linux packages & CPack on Unix-like systems &
\depcode{dpkg-shlibdeps} enables DEB generation; \depcode{rpmbuild} enables RPM
generation for non-parallel builds. \\
\hline
\end{longtable}
\endgroup

\subsubsection{External Source Defaults}

The following default upstream source locations are encoded in
\depcode{config/CacheURLs.cmake} for dependencies that HDF5 can fetch or whose
values are passed to external filter-plugin builds.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.21\linewidth}p{0.19\linewidth}p{0.51\linewidth}@{}}
\hline
\textbf{Dependency} & \textbf{Default version/tag} & \textbf{Default source} \\
\hline
zlib & \depcode{1.3.2} / \depcode{v1.3.2} &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.2/zlib-1.3.2.tar.gz} \\
zlib-ng & \depcode{2.3.3} &
\url{https://github.com/zlib-ng/zlib-ng.git} or
\url{https://github.com/zlib-ng/zlib-ng/archive/refs/tags/2.3.3.tar.gz} \\
libaec & \depcode{1.1.6} / \depcode{v1.1.6} &
\url{https://github.com/MathisRosenhauer/libaec.git} or
\url{https://github.com/MathisRosenhauer/libaec/releases/download/v1.1.6/libaec-1.1.6.tar.gz} \\
HDF5 filter plugins & \depcode{master} &
\url{https://github.com/HDFGroup/hdf5_plugins.git} or
\url{https://github.com/HDFGroup/hdf5_plugins/releases/download/snapshot/hdf5_plugins-master.tar.gz} \\
KWSys API-test helper & \depcode{master} tarball &
\url{https://gitlab.kitware.com/utils/kwsys/-/archive/master/kwsys-master.tar.gz} \\
h5py Python examples & \depcode{master} &
\url{https://github.com/h5py/h5py.git} \\
bitshuffle & \depcode{master} or \depcode{0.5.2} tarball &
\url{https://github.com/kiyo-masui/bitshuffle.git} or
\url{https://github.com/kiyo-masui/bitshuffle/archive/refs/tags/bitshuffle-0.5.2.tar.gz} \\
Blosc & \depcode{1.21.6} &
\url{https://github.com/Blosc/c-blosc.git} or
\url{https://github.com/Blosc/c-blosc/archive/refs/tags/c-blosc-1.21.6.tar.gz} \\
Blosc zlib helper & \depcode{develop} or \depcode{1.3.1} tarball &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.1/zlib-1.3.1.tar.gz} \\
Blosc2 & \depcode{2.17.1} &
\url{https://github.com/Blosc/c-blosc2.git} or
\url{https://github.com/Blosc/c-blosc2/archive/refs/tags/c-blosc2-2.17.1.tar.gz} \\
Blosc2 zlib helper & \depcode{develop} or \depcode{1.3.1} tarball &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.1/zlib-1.3.1.tar.gz} \\
bzip2 & \depcode{bzip2-1.0.8} &
\url{https://github.com/libarchive/bzip2.git} or
\url{https://github.com/libarchive/bzip2/archive/refs/tags/bzip2-bzip2-1.0.8.tar.gz} \\
fpzip & \depcode{develop} or \depcode{1.3.0} tarball &
\url{https://github.com/LLNL/fpzip.git} or
\url{https://github.com/LLNL/fpzip/releases/download/1.3.0/fpzip-1.3.0.tar.gz} \\
JPEG & \depcode{jpeg-9e} &
\url{https://github.com/libjpeg-turbo/libjpeg-turbo.git} or
\url{https://www.ijg.org/files/jpegsrc.v9e.tar.gz} \\
LZ4 & \depcode{dev} or \depcode{1.10.0} tarball &
\url{https://github.com/lz4/lz4.git} or
\url{https://github.com/lz4/lz4/releases/download/v1.10.0/lz4-1.10.0.tar.gz} \\
LZF & \depcode{3.6} tarball &
\url{http://dist.schmorp.de/liblzf/liblzf-3.6.tar.gz} \\
SZ & \depcode{master} or \depcode{2.1.12.5} tarball &
\url{https://github.com/szcompressor/SZ.git} or
\url{https://github.com/szcompressor/SZ/releases/download/v2.1.12.5/SZ-2.1.12.5.tar.gz} \\
ZFP & \depcode{develop} or \depcode{1.0.1} tarball &
\url{https://github.com/LLNL/zfp.git} or
\url{https://github.com/LLNL/zfp/releases/download/1.0.0/zfp-1.0.1.tar.gz} \\
Zstandard & \depcode{1.5.7} &
\url{https://github.com/facebook/zstd.git} or
\url{https://github.com/facebook/zstd/releases/download/v1.5.7/zstd-1.5.7.tar.gz} \\
\hline
\end{longtable}
\endgroup

\subsection{Downstream Dependencies}

For this audit, downstream dependencies are projects that compile against,
dynamically load, package, or define data conventions on top of HDF5. The HDF5
ecosystem is much larger than any static list, so the most important downstreams
are best identified by blast radius: broad installation footprint, long-lived
archival data, domain-standard file conventions, or use as an intermediate layer
for other tools.

\subsubsection{High-Impact Downstream Families}

\begin{description}[leftmargin=1.5em,style=nextline]
    \item[Scientific data models and exchange formats.]
    The highest-impact downstream is netCDF-4, which uses HDF5 as the storage
    layer for the enhanced netCDF data model and is central to climate, weather,
    ocean, atmospheric, and other Earth-system data workflows~\cite{netcdf4-format}.
    NASA HDF-EOS5, NeXus, and CGNS similarly define domain-specific conventions
    over HDF5 for Earth observation, neutron/x-ray/muon facilities, and CFD or
    aerospace simulation data~\cite{hdf-eos5,nexus-format,cgns-hdf5}. For these
    formats, HDF5 compatibility is not just a library question: it affects the
    readability of long-lived archives and the ability of independent tools to
    interpret a shared schema.

    \item[Language bindings and analysis libraries.]
    h5py, PyTables, pandas HDFStore, Bioconductor's rhdf5, HDF5.jl, and MATLAB's
    HDF5 interfaces make HDF5 a routine dependency for Python, R, Julia, and
    MATLAB users~\cite{h5py-docs,pytables-docs,pandas-hdfstore,rhdf5,hdf5-jl,matlab-hdf5}.
    These packages are important because they often sit at the boundary between
    upstream HDF5 builds and end-user scripts. Their packaging choices determine
    whether users load a bundled HDF5, a system HDF5, or an environment-provided
    HDF5, and therefore whether ABI, plugin-path, thread-safety, MPI, and filter
    support match the files being opened.

    \item[Geospatial, visualization, and inspection tools.]
    GDAL's HDF5 driver, ParaView/VTK-family builds, ITK's HDF5 module, HDFView,
    and similar readers expose HDF5 files to interactive and automated analysis
    workflows~\cite{gdal-hdf5,paraview-build,itk-hdf5}. This is an important
    security boundary: these tools are commonly used to inspect third-party data
    files, so robustness against malformed HDF5 objects, unusual filters, and
    unexpected VFD/plugin configuration has direct operational value.

    \item[HPC and parallel-I/O frameworks.]
    ADIOS2 includes an HDF5 engine, while netCDF-4, CGNS, h5py, and many
    simulation applications can be built against serial or parallel HDF5
    variants~\cite{adios2-hdf5}. These downstreams amplify configuration risk:
    MPI ABI compatibility, collective I/O semantics, file locking behavior,
    subfiling support, and optional VFDs must line up across the application,
    HDF5, MPI, and filesystem stack.

    \item[Machine-learning and model-artifact workflows.]
    TensorFlow/Keras still supports older HDF5 model and weight files even as
    newer native formats are preferred~\cite{tensorflow-keras-hdf5}. This keeps
    HDF5 in the model-exchange and training-checkpoint ecosystem, usually through
    h5py, and makes HDF5 parser behavior relevant for model files obtained from
    external sources.
\end{description}

\subsubsection{Representative Downstream Projects}

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}
\begin{longtable}{@{}p{0.24\linewidth}p{0.32\linewidth}p{0.35\linewidth}@{}}
\hline
\textbf{Downstream} & \textbf{Role} & \textbf{Supply-chain relevance} \\
\hline
netCDF-4 &
Common scientific data model and format layered on HDF5. &
Large transitive footprint through climate, weather, ocean, remote-sensing, and
HPC data stacks. Uses a constrained HDF5 subset, so compatibility depends on both
HDF5 behavior and netCDF conventions. \\
HDF-EOS5 &
NASA Earth-observation format built on HDF5 concepts and libraries. &
Long-lived public archives and many third-party readers make backward
compatibility and robust metadata handling high-value. \\
NeXus &
Facility data convention for neutron, x-ray, and muon science. &
Uses HDF5 groups, datasets, and attributes as a standardized experimental-data
container, so HDF5 schema interpretation affects reproducibility across
facilities. \\
CGNS &
CFD and aerospace data standard with an HDF5 mapping. &
Important for mesh and simulation-result exchange; parallel-HDF5 behavior and
large-file support affect production simulation workflows. \\
h5py, PyTables, and pandas HDFStore &
Python interfaces and tabular/array storage layers. &
Common bridge between binary HDF5 files and Python analysis. Wheel and conda
packaging can hide which HDF5 build is actually loaded. \\
rhdf5, HDF5.jl, MATLAB HDF5 APIs &
R, Julia, and MATLAB access layers. &
Expose HDF5 to laboratory and scientific-computing workflows where files are
often exchanged outside controlled build environments. \\
GDAL and geospatial readers &
Raster and subdataset access for HDF5-based products. &
Converts HDF5 issues into geospatial ingestion issues, especially for
remote-sensing products with complex metadata and nested subdatasets. \\
ParaView/VTK, ITK, HDFView, and viewers &
Interactive inspection, visualization, and conversion. &
Frequently open untrusted or externally generated files; defensive parsing,
filter availability, and plugin search paths matter. \\
ADIOS2 and simulation I/O layers &
Alternative I/O stacks with HDF5 read/write engines or compatibility paths. &
Parallel and high-throughput workflows depend on matching MPI, filesystem, and
HDF5 feature assumptions across the whole stack. \\
TensorFlow/Keras HDF5 artifacts &
Legacy model and weight serialization. &
Keeps HDF5 in ML artifact exchange; malformed or incompatible model files
surface through h5py and HDF5 library behavior. \\
\hline
\end{longtable}
\endgroup

\subsubsection{Audit Implications}

\begin{itemize}
    \item HDF5 changes have a large transitive blast radius because many users
    reach HDF5 through netCDF, GDAL, h5py, MATLAB, or domain formats rather than
    through the HDF5 C API directly.
    \item ABI and packaging mismatches are a recurring downstream risk. Python,
    R, Julia, MATLAB, MPI, and system-package environments may all load different
    HDF5 builds in the same organization.
    \item File parsing is a practical security boundary. Viewers, converters,
    notebooks, and data-ingestion jobs often process files received from outside
    the organization that built the HDF5 library.
    \item Optional features can become hidden interoperability requirements:
    filter plugins, SZIP/zlib variants, ROS3/HDFS/VFD support, parallel HDF5,
    and file-locking behavior may be assumed by downstream data even when they
    are not present in a default HDF5 build.
    \item Long-lived scientific archives make deprecation slower than ordinary
    application software. Even when a downstream project prefers a newer format,
    historic HDF5 files remain operationally relevant.
\end{itemize}

\subsection{Audit Findings and Mitigation Priorities}

\end{document}
