\documentclass[../main.tex]{subfiles}

\begin{document}

%% ======================================================================
%%  Appendix 1 – ???
%% ======================================================================
\section{Software Supply Chain Details}
\label{app:supply-chain}


The \code{HDF5_ALLOW_EXTERNAL_SUPPORT} modes apply to filter-related
dependencies. Some test and example configurations have their own source-fetch
paths, such as KWSys for the API test driver and h5py for Python examples. MPI,
HDFS, ROS3, OpenSSL, Java, documentation, packaging, and developer tools must be
provided by the build environment.

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.16\widetablewidth}p{0.38\widetablewidth}p{0.42\widetablewidth}@{}}
\caption{Dependency source modes}
\label{tab:dependency-source-modes}\\
\hline
\textbf{Mode} & \textbf{Relevant options} & \textbf{Meaning} \\
\hline
System packages & \code{HDF5_ALLOW_EXTERNAL_SUPPORT=NO} &
CMake uses \code{find_package()}, \code{find_library()}, or
\code{find_program()} to locate installed dependencies. This is the default. \\
Git fetch & \code{HDF5_ALLOW_EXTERNAL_SUPPORT=GIT} plus a
dependency-specific \code{*_USE_EXTERNAL=ON} option &
CMake downloads source from the configured \code{*_GIT_URL} and
\code{*_GIT_TAG} values and builds it with HDF5. \\
Tarball fetch & \code{HDF5_ALLOW_EXTERNAL_SUPPORT=TGZ} plus a
dependency-specific \code{*_USE_EXTERNAL=ON} option &
CMake downloads from \code{*_TGZ_ORIGPATH} and \code{*_TGZ_NAME}, or uses
\code{TGZPATH} when the dependency-specific \code{*_USE_LOCALCONTENT=ON}
option is set. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.3\widetablewidth}p{0.22\widetablewidth}p{0.45\widetablewidth}@{}}
\caption{Baseline build}
\label{tab:baseline}\\
\hline
\textbf{Dependency} & \textbf{Trigger} & \textbf{Notes} \\
\hline
CMake 3.26 or newer & Always &
Required by the top-level \code{cmake_minimum_required()}. \\
C compiler and native build tool & Always &
The top-level project is a C project. \\
Platform C library, headers, and system calls & Always &
CMake probes standard headers, types, functions, byte order, large-file support,
file locking, \code{pread}/\code{pwrite}, and related platform features. \\
Platform link libraries & Platform-dependent &
CMake adds libraries only when probes show they are needed: \code{m},
\code{dl}, \code{rt}, \code{posix4}, \code{ucb}, \code{ws2_32},
\code{wsock32}, and \code{shlwapi}. \\
PkgConfig & Optional probe &
Used only when available. It can help some find modules, such as zlib-ng and
libaec, but it is not a hard requirement for the default build. \\
Threads & Found in every configure; required by thread features &
CMake searches for C11 threads, Win32 threads, or pthreads. When found, HDF5
records thread support and links \code{Threads::Threads};
\code{HDF5_ENABLE_THREADSAFE}, \code{HDF5_ENABLE_CONCURRENCY}, and
\code{HDF5_ENABLE_SUBFILING_VFD} require it. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.20\widetablewidth}p{0.27\widetablewidth}p{0.20\widetablewidth}p{0.24\widetablewidth}@{}}
\caption{HDF5 feature triggers}
\label{tab:feature-triggers}\\
\hline
\textbf{Feature or component} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Parallel HDF5 & \code{HDF5_ENABLE_PARALLEL=ON} &
MPI C with MPI-IO, MPI standard 3.0 or newer &
Adds \code{MPI::MPI_C} to public link libraries. Parallel filtered writes and
large parallel I/O are enabled only when the required MPI symbols are present.
Parallel tests also need an MPI launcher such as \code{mpiexec}. \\
Parallel Fortran & \code{HDF5_ENABLE_PARALLEL=ON},
\code{HDF5_BUILD_FORTRAN=ON} &
MPI Fortran &
The Fortran build adds MPI Fortran libraries. The \code{mpi_f08} module is
used when available. \\
Thread-safe library & \code{HDF5_ENABLE_THREADSAFE=ON} &
Threads &
Unsupported with parallel, Fortran, C++, or high-level library unless
\code{HDF5_ALLOW_UNSUPPORTED=ON}. On Windows, static libraries are rejected. \\
Multi-threaded concurrency & \code{HDF5_ENABLE_CONCURRENCY=ON} &
Threads &
Experimental. Mutually exclusive with \code{HDF5_ENABLE_THREADSAFE}; has the
same unsupported combinations as thread-safety unless overridden with
\code{HDF5_ALLOW_UNSUPPORTED=ON}. \\
C++ wrappers & \code{HDF5_BUILD_CPP_LIB=ON} &
C++ compiler &
Unsupported with parallel unless \code{HDF5_ALLOW_UNSUPPORTED=ON}.
High-level C++ wrappers also require \code{HDF5_BUILD_HL_LIB=ON}. \\
Fortran wrappers & \code{HDF5_BUILD_FORTRAN=ON} &
Fortran compiler with Fortran 2003 \code{ISO_C_BINDING} support &
Unsupported with thread-safety or concurrency unless
\code{HDF5_ALLOW_UNSUPPORTED=ON}. High-level Fortran wrappers also require
\code{HDF5_BUILD_HL_LIB=ON}. \\
Java wrappers & \code{HDF5_BUILD_JAVA=ON} &
Java/JDK, shared HDF5 libraries, Java dependency JARs &
Java requires shared HDF5 libraries. Java 11 or newer is required for the JNI
implementation. Java tests additionally require the JUnit and Hamcrest JARs. \\
Java JNI implementation & \code{HDF5_BUILD_JAVA=ON},
\code{HDF5_ENABLE_JNI=ON}, or any Java version earlier than 25 &
JNI headers and libraries &
JNI is the default Java implementation. HDFS also requires JNI. \\
Java FFM implementation & \code{HDF5_BUILD_JAVA=ON}, Java 25 or newer,
\code{HDF5_ENABLE_JNI=OFF} &
\code{jextract} from \code{JEXTRACT_HOME/bin} or \code{JAVA_HOME/bin} &
Generates FFM bindings at configure time. \\
Java logging/test JARs & \code{HDF5_BUILD_JAVA=ON} &
\code{slf4j-api}, \code{slf4j-nop}, \code{slf4j-simple}; plus JUnit 4
and Hamcrest for tests &
The default paths point to bundled JARs under \code{java/lib}. They can be
overridden with \code{HDF5_JAVA_*_JAR} cache variables.
\code{HDF5_ENABLE_MAVEN_DEPLOY=ON} generates Maven-style artifacts and POM
files; the CMake build does not locate or run \code{mvn}. \\
Documentation & \code{HDF5_BUILD_DOC=ON} &
Doxygen &
\code{HDF5_ENABLE_DOXY_WARNINGS=ON} makes Doxygen warnings fail the build. \\
Header regeneration & \code{HDF5_GENERATE_HEADERS=ON} &
Perl &
Perl is optional otherwise. If it is missing, generated headers are not
regenerated. \\
Python examples & \code{H5EXAMPLE_BUILD_PYTHON=ON} &
Python3 interpreter, Python development files, NumPy, h5py source, pip &
The examples CMake project fetches \code{h5py} from Git and installs it with
\texttt{python3 -m pip}. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.19\widetablewidth}p{0.28\widetablewidth}p{0.22\widetablewidth}p{0.22\widetablewidth}@{}}
\caption{HDF5 filter dependency choices}
\label{tab:filter-dependencies}\\
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Source choices} \\
\hline
DEFLATE filter & \code{HDF5_ENABLE_ZLIB_SUPPORT=ON},
\code{HDF5_USE_ZLIB_NG=OFF} &
zlib &
System package by default; \code{ZLIB_USE_EXTERNAL=ON} can build zlib from
Git or TGZ. \code{HDF5_USE_ZLIB_STATIC=ON} prefers a static library. \\
DEFLATE filter through zlib-ng & \code{HDF5_ENABLE_ZLIB_SUPPORT=ON},
\code{HDF5_USE_ZLIB_NG=ON} &
zlib-ng &
System package by default; \code{ZLIB_USE_EXTERNAL=ON} can build zlib-ng from
Git or TGZ. zlib compatibility mode is detected. \\
SZIP filter & \code{HDF5_ENABLE_SZIP_SUPPORT=ON} &
libaec with libsz compatibility, or another package selected by
\code{LIBAEC_PACKAGE_NAME} &
System package by default; \code{SZIP_USE_EXTERNAL=ON} can build libaec from
Git or TGZ. \code{HDF5_USE_LIBAEC_STATIC=ON} prefers static libraries. \\
HDF5 filter plugins project & \code{HDF5_ENABLE_PLUGIN_SUPPORT=ON} &
Installed \code{h5pl}/HDF5 filter plugins package, or the
\code{hdf5_plugins} project &
\code{PLUGIN_USE_EXTERNAL=ON} can build the plugin project from Git or TGZ.
Plugin support is disabled for the 1.6 API. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.19\widetablewidth}p{0.28\widetablewidth}p{0.22\widetablewidth}p{0.22\widetablewidth}@{}}
\caption{HDF5 VFD and VOL connector dependency choices}
\label{tab:vfd-vol-dependency-choices}\\
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Direct VFD & \code{HDF5_ENABLE_DIRECT_VFD=ON} &
POSIX \code{O_DIRECT} and \code{posix_memalign()} support &
No third-party library. Configuration fails if the platform support is missing. \\
Mirror VFD & \code{HDF5_ENABLE_MIRROR_VFD=ON} &
POSIX socket/fork headers and functions &
No third-party library. Requires \code{netinet/in.h}, \code{netdb.h},
\code{arpa/inet.h}, \code{sys/socket.h}, and \code{fork()}. \\
ROS3 VFD & \code{HDF5_ENABLE_ROS3_VFD=ON} &
\code{aws-c-s3} CMake package &
HDF5 does not fetch this library. Building \code{aws-c-s3} from source can
also require AWS C libraries such as \code{aws-c-common},
\code{aws-checksums}, \code{aws-c-cal}, \code{aws-c-io},
\code{aws-c-compression}, \code{aws-c-http}, \code{aws-c-sdkutils},
\code{aws-c-auth}, and, on Linux, \code{aws-lc} and \code{s2n-tls}.
ROS3 docker-proxy tests additionally require \code{docker} and the AWS CLI. \\
HDFS VFD & \code{HDF5_ENABLE_HDFS=ON} &
JNI/JVM and Hadoop \code{libhdfs} &
\code{HADOOP_HOME} is used to locate \code{hdfs.h}, \code{libhdfs},
and the Hadoop version command. Non-MSVC builds also add \code{-pthread}. \\
Subfiling VFD and IOC VFD & \code{HDF5_ENABLE_PARALLEL=ON},
\code{HDF5_ENABLE_SUBFILING_VFD=ON} &
MPI C with \code{MPI_Comm_split_type}, plus Threads &
Not available on Windows. IOC VFD is built when subfiling is enabled. \\
External VOL connectors & \code{HDF5_VOL_ALLOW_EXTERNAL=GIT} or
\code{LOCAL_DIR} with \code{HDF5_VOL_URLnn} or \code{HDF5_VOL_PATHnn} &
Connector-defined &
Up to ten CMake-based VOL connectors can be added. HDF5 patches connector CMake
code to use the in-tree HDF5 build, but each connector may bring its own
dependencies. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.21\widetablewidth}p{0.29\widetablewidth}p{0.20\widetablewidth}p{0.21\widetablewidth}@{}}
\caption{Tools, Tests, and Security Features}
\label{tab:tools-tests-security-features}\\
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Signed plugin verification & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON} &
OpenSSL libcrypto 1.1.0 or newer &
Adds \code{OpenSSL::Crypto} to HDF5. \code{HDF5_LOCK_PLUGIN_KEYSTORE=ON}
also requires \code{HDF5_PLUGIN_KEYSTORE_DIR}. \\
\code{h5sign} tool & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\code{HDF5_BUILD_TOOLS=ON} &
OpenSSL libcrypto &
Used to sign HDF5 plugins. \\
Signed-plugin tests & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\code{BUILD_TESTING=ON} &
\code{openssl} executable &
Tests generate RSA keys with the OpenSSL command-line tool. \\
Parallel tools utilities & \code{HDF5_ENABLE_PARALLEL=ON},
\code{HDF5_BUILD_PARALLEL_TOOLS=ON} &
MFU, CIRCLE, DTCMP &
CMake looks for headers and libraries, using \code{MFU_ROOT} as a hint. \\
API test driver & \code{BUILD_TESTING=ON},
\code{HDF5_TEST_API_ENABLE_DRIVER=ON} &
KWSys &
The test driver fetches KWSys from a tarball or from \code{TGZPATH} when
\code{KWSYS_USE_LOCALCONTENT=ON}. \\
Shell-script tests & \code{BUILD_TESTING=ON} &
PowerShell or Bash &
CMake uses \code{pwsh}/\code{powershell} when available; otherwise Unix
builds look for \code{bash}. \\
ROS3 VFD tests & \code{HDF5_ENABLE_ROS3_VFD=ON},
\code{HDF5_ENABLE_ROS3_VFD_DOCKER_PROXY=ON} &
Docker daemon and AWS CLI &
Used to run S3 proxy tests. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.21\widetablewidth}p{0.29\widetablewidth}p{0.20\widetablewidth}p{0.21\widetablewidth}@{}}
\caption{Tools, Tests, and Security Features}
\label{tab:tools-tests-security-features}\\
\hline
\textbf{Feature} & \textbf{Build options} & \textbf{Dependency} &
\textbf{Notes} \\
\hline
Signed plugin verification & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON} &
OpenSSL libcrypto 1.1.0 or newer &
Adds \code{OpenSSL::Crypto} to HDF5. \code{HDF5_LOCK_PLUGIN_KEYSTORE=ON}
also requires \code{HDF5_PLUGIN_KEYSTORE_DIR}. \\
\code{h5sign} tool & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\code{HDF5_BUILD_TOOLS=ON} &
OpenSSL libcrypto &
Used to sign HDF5 plugins. \\
Signed-plugin tests & \code{HDF5_REQUIRE_SIGNED_PLUGINS=ON},
\code{BUILD_TESTING=ON} &
\code{openssl} executable &
Tests generate RSA keys with the OpenSSL command-line tool. \\
Parallel tools utilities & \code{HDF5_ENABLE_PARALLEL=ON},
\code{HDF5_BUILD_PARALLEL_TOOLS=ON} &
MFU, CIRCLE, DTCMP &
CMake looks for headers and libraries, using \code{MFU_ROOT} as a hint. \\
API test driver & \code{BUILD_TESTING=ON},
\code{HDF5_TEST_API_ENABLE_DRIVER=ON} &
KWSys &
The test driver fetches KWSys from a tarball or from \code{TGZPATH} when
\code{KWSYS_USE_LOCALCONTENT=ON}. \\
Shell-script tests & \code{BUILD_TESTING=ON} &
PowerShell or Bash &
CMake uses \code{pwsh}/\code{powershell} when available; otherwise Unix
builds look for \code{bash}. \\
ROS3 VFD tests & \code{HDF5_ENABLE_ROS3_VFD=ON},
\code{HDF5_ENABLE_ROS3_VFD_DOCKER_PROXY=ON} &
Docker daemon and AWS CLI &
Used to run S3 proxy tests. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.25\widetablewidth}p{0.34\widetablewidth}p{0.32\widetablewidth}@{}}
\caption{Developer and Packaging Dependencies}
\label{tab:developer-packaging-dependencies}\\
\hline
\textbf{Configuration} & \textbf{Build options} & \textbf{Dependency} \\
\hline
Analyzer tooling & \code{HDF5_ENABLE_ANALYZER_TOOLS=ON} &
\code{clang-tidy}, \code{include-what-you-use}, and \code{cppcheck}
are probed and used by the analyzer macros. \\
Source formatting targets & \code{HDF5_ENABLE_FORMATTERS=ON} &
\code{clang-format} and \code{cmake-format}. \\
Code coverage with Clang/IntelLLVM & \code{HDF5_ENABLE_COVERAGE=ON},
\code{CODE_COVERAGE=ON} &
\code{llvm-cov} and \code{llvm-profdata}; AppleClang uses
\texttt{xcrun llvm-cov} and \texttt{xcrun llvm-profdata}. \\
Code coverage with GCC & \code{HDF5_ENABLE_COVERAGE=ON},
\code{CODE_COVERAGE=ON} &
\code{lcov} and \code{genhtml}; fallback instrumentation can also link
\code{gcov}. \\
Sanitizers & \code{HDF5_ENABLE_SANITIZERS=ON},
\code{HDF5_USE_SANITIZER=<kind>} &
Compiler sanitizer runtime support. Supported values depend on compiler and
platform. \\
AFL support module & Explicit inclusion of
\code{config/sanitizer/afl-fuzzing.cmake} &
\code{afl-cc} and \code{afl-c++}. This module is available but is not
included by the top-level HDF5 build by default. \\
Dependency graph support module & Explicit inclusion of
\code{config/sanitizer/dependency-graph.cmake} &
Graphviz \code{dot}. This module is available but is not included by the
top-level HDF5 build by default. \\
Windows installers & CPack on Windows &
NSIS and WiX are probed when available. \\
Linux packages & CPack on Unix-like systems &
\code{dpkg-shlibdeps} enables DEB generation; \code{rpmbuild} enables RPM
generation for non-parallel builds. \\
\hline
\end{longtable}
\endgroup

\begingroup
\small
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.15}

\setlength{\LTleft}{-0.7in}
\setlength{\LTright}{-0.7in}

\begin{longtable}{@{}p{0.21\widetablewidth}p{0.19\widetablewidth}p{0.51\widetablewidth}@{}}
\caption{External Source Defaults}
\label{tab:external-source-defaults}\\
\hline
\textbf{Dependency} & \textbf{Default version/tag} & \textbf{Default source} \\
\hline
zlib & \code{1.3.2} / \code{v1.3.2} &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.2/zlib-1.3.2.tar.gz} \\
zlib-ng & \code{2.3.3} &
\url{https://github.com/zlib-ng/zlib-ng.git} or
\url{https://github.com/zlib-ng/zlib-ng/archive/refs/tags/2.3.3.tar.gz} \\
libaec & \code{1.1.6} / \code{v1.1.6} &
\url{https://github.com/MathisRosenhauer/libaec.git} or
\url{https://github.com/MathisRosenhauer/libaec/releases/download/v1.1.6/libaec-1.1.6.tar.gz} \\
HDF5 filter plugins & \code{master} &
\url{https://github.com/HDFGroup/hdf5_plugins.git} or
\url{https://github.com/HDFGroup/hdf5_plugins/releases/download/snapshot/hdf5_plugins-master.tar.gz} \\
KWSys API-test helper & \code{master} tarball &
\url{https://gitlab.kitware.com/utils/kwsys/-/archive/master/kwsys-master.tar.gz} \\
h5py Python examples & \code{master} &
\url{https://github.com/h5py/h5py.git} \\
bitshuffle & \code{master} or \code{0.5.2} tarball &
\url{https://github.com/kiyo-masui/bitshuffle.git} or
\url{https://github.com/kiyo-masui/bitshuffle/archive/refs/tags/bitshuffle-0.5.2.tar.gz} \\
Blosc & \code{1.21.6} &
\url{https://github.com/Blosc/c-blosc.git} or
\url{https://github.com/Blosc/c-blosc/archive/refs/tags/c-blosc-1.21.6.tar.gz} \\
Blosc zlib helper & \code{develop} or \code{1.3.1} tarball &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.1/zlib-1.3.1.tar.gz} \\
Blosc2 & \code{2.17.1} &
\url{https://github.com/Blosc/c-blosc2.git} or
\url{https://github.com/Blosc/c-blosc2/archive/refs/tags/c-blosc2-2.17.1.tar.gz} \\
Blosc2 zlib helper & \code{develop} or \code{1.3.1} tarball &
\url{https://github.com/madler/zlib.git} or
\url{https://github.com/madler/zlib/releases/download/v1.3.1/zlib-1.3.1.tar.gz} \\
bzip2 & \code{bzip2-1.0.8} &
\url{https://github.com/libarchive/bzip2.git} or
\url{https://github.com/libarchive/bzip2/archive/refs/tags/bzip2-bzip2-1.0.8.tar.gz} \\
fpzip & \code{develop} or \code{1.3.0} tarball &
\url{https://github.com/LLNL/fpzip.git} or
\url{https://github.com/LLNL/fpzip/releases/download/1.3.0/fpzip-1.3.0.tar.gz} \\
JPEG & \code{jpeg-9e} &
\url{https://github.com/libjpeg-turbo/libjpeg-turbo.git} or
\url{https://www.ijg.org/files/jpegsrc.v9e.tar.gz} \\
LZ4 & \code{dev} or \code{1.10.0} tarball &
\url{https://github.com/lz4/lz4.git} or
\url{https://github.com/lz4/lz4/releases/download/v1.10.0/lz4-1.10.0.tar.gz} \\
LZF & \code{3.6} tarball &
\url{http://dist.schmorp.de/liblzf/liblzf-3.6.tar.gz} \\
SZ & \code{master} or \code{2.1.12.5} tarball &
\url{https://github.com/szcompressor/SZ.git} or
\url{https://github.com/szcompressor/SZ/releases/download/v2.1.12.5/SZ-2.1.12.5.tar.gz} \\
ZFP & \code{develop} or \code{1.0.1} tarball &
\url{https://github.com/LLNL/zfp.git} or
\url{https://github.com/LLNL/zfp/releases/download/1.0.0/zfp-1.0.1.tar.gz} \\
Zstandard & \code{1.5.7} &
\url{https://github.com/facebook/zstd.git} or
\url{https://github.com/facebook/zstd/releases/download/v1.5.7/zstd-1.5.7.tar.gz} \\
\hline
\end{longtable}
\endgroup

\end{document}
